### 2022-05-10 analysis/gafo-cc-analysis.Rmd
add first round of data

---
title: "Gafo-cc-analysis"
output: html_document
date: '2022-03-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brms)
```

```{r}
mdata <- read_csv("../data/gafo-cc-clean-data.csv")%>%
  mutate(ageinyears = scale(ageinyears),
         targetcentralityx = scale(targetcentralityx),
         trialnr = scale(trialnr),
         children = scale(children),
         youngerchildren = scale(youngerchildren),
         touchscreen = factor(touchscreen.x),
        screen = factor(screen))
```

```{r}
bm_base <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_null <- brm(abs(clickdistfromtargetcenterx) ~ 1 + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

loo_compare(bm_base, bm_null)%>%as_tibble(rownames = "model")

```

```{r}
plot(bm_age)

pp_check(bm_age)+xlim(0,2000)

bm_age
```

```{r}
bm_screen <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + screen + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_touch <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + touchscreen + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))
```

```{r}
bm_child <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + children + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_young <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + youngerchildren + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))
```