### 2022-06-21 analysis/gafo-cc-analysis.Rmd
prep model run on cluster

---
title: "Gafo-cc-analysis"
output: html_document
date: '2022-03-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brms)
library(tidybayes)
```

```{r}
mdata <- read_csv("../data/gafo-cc-clean-data.csv")%>%
  mutate(ageinyears = scale(ageinyears),
         targetcentralityx = scale(targetcentralityx),
         trialnr = scale(trialnr),
         children = scale(children),
         household = scale(household),
         younger_children = scale(younger_children),
         touchscreen = factor(touchscreen),
        screen = factor(screen))%>%
  filter(community == "leipzig" |community == "akure")

mdata%>%
  ggplot(aes(x = abs(clickdistfromtargetcenterx)))+
  geom_density()

c(
  lower = exp(4 - 2 * 1.75),
  higher = exp(4 + 2 * 1.75)
)

```

```{r}
bm_gam <- brm(abs(clickdistfromtargetcenterx) ~ 1, 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_log <- brm(abs(clickdistfromtargetcenterx) ~ 1, 
                              data   = mdata, 
                              family = lognormal(),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_slog <- brm(abs(clickdistfromtargetcenterx) ~ 1, 
                              data   = mdata, 
                              family = shifted_lognormal(),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_exg<- brm(abs(clickdistfromtargetcenterx) ~ 1, 
                              data   = mdata, 
                              family = exgaussian(),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_wei<- brm(abs(clickdistfromtargetcenterx) ~ 1, 
                              data   = mdata, 
                              family = weibull(),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

loo_compare(bm_gam, bm_log,bm_slog, bm_exg,bm_wei)%>%as_tibble(rownames = "model")

pp_check(bm_gam)+xlim(0,3000)
pp_check(bm_log)+xlim(0,3000)
pp_check(bm_slog)+xlim(0,3000)
pp_check(bm_exg)+xlim(0,3000)
pp_check(bm_wei)+xlim(0,3000)
```

```{r}
bm_base_gam <- brm(abs(clickdistfromtargetcenterx) ~ targetcentralityx + (trialnr|subjid) + (0+ageinyears|community), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 2000, 
                              iter   = 5000,
                   control = list(adapt_delta = 0.95),
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_base_wei <- brm(abs(clickdistfromtargetcenterx) ~ targetcentralityx + (trialnr|subjid) + (0+ageinyears|community), 
                              data   = mdata, 
                              family = weibull(),
                              warmup = 2000, 
                              iter   = 5000, 
                   control = list(adapt_delta = 0.95),
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

loo_compare(bm_base_gam,bm_base_wei)%>%as_tibble(rownames = "model")

pp_check(bm_base_gam)+xlim(0,3000)
pp_check(bm_base_wei)+xlim(0,3000)

```

```{r}
bm_base <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + (trialnr|subjid) + (ageinyears|community), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_null <- brm(abs(clickdistfromtargetcenterx) ~ 1 + (trialnr|subjid)+ (1|community), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

loo_compare(bm_base, bm_null)%>%as_tibble(rownames = "model")

```

```{r}
plot(bm_base)

pp_check(bm_base)

bm_age
```

```{r}
ps <- bm_base%>%posterior_samples()%>%
  select(starts_with("r_community"))%>%
  pivot_longer(cols = everything(), names_to = "param", values_to = "value")%>%
  filter(grepl("Intercept", param))

ggplot(ps,aes(y = param, x = value)) +
 stat_halfeye(alpha = .7, .width = c(.95, .80)) +
    labs(x = "Posterior estimate", y ="")+
  guides(fill = "none")+
  #geom_vline(xintercept = 0, linetype = "dashed") +
  xlim(-0.2, 0.5)+
  theme_minimal()
```

```{r}
bm_screen <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx  + (1|subjid) + (screen + children + household|community), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_touch <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + (1|subjid) + (1|community), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

```
```{r}
ps <- bm_screen%>%posterior_samples()%>%
  select(starts_with("r_community"))%>%
  pivot_longer(cols = everything(), names_to = "param", values_to = "value")%>%
  filter(grepl("Intercept", param))

ggplot(ps,aes(y = param, x = value)) +
 stat_halfeye(alpha = .7, .width = c(.95, .80)) +
    labs(x = "Posterior estimate", y ="")+
  guides(fill = "none")+
  #geom_vline(xintercept = 0, linetype = "dashed") +
  xlim(-0.2, 0.5)+
  theme_minimal()
```

```{r}
bm_child <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + children + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))

bm_young <- brm(abs(clickdistfromtargetcenterx) ~ ageinyears  + targetcentralityx + youngerchildren + (trialnr|subjid), 
                              data   = mdata, 
                              family = Gamma(link = "log"),
                              warmup = 1000, 
                              iter   = 4000, 
                              chains = 4, 
                              cores  = 4)%>%
  add_criterion(., c("waic","loo"))
```