### 2022-07-08 analysis/gafo-cc-visuals.Rmd
run cc model leip and ak on cluster

---
title: "Gafo-cc visuals"
output: html_document
---

```{r}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(tidyboot)
```

```{r}
data <- read_csv("../data/gafo-cc-clean-data.csv")
```

```{r}
data%>%
  group_by(community)%>%
  summarise(n= n_distinct(subjid))

data%>%
  summarise(n= n_distinct(subjid))
```

```{r}
data%>%
  group_by(community,subjid, ageinyears)%>%
  summarise(impr = mean(abs(clickdistfromtargetcenterx)))%>%
  ggplot(aes(x = ageinyears, y = impr, col = community))+
  geom_point(pch = 1)+
  geom_smooth(method = "glm")+
  scale_color_colorblind(name = "", labels = c("Nigeria", "Namibia", "Sambia", "Namibia", "Germany"))+
  ylim(0,1000)+
  labs(x = "Age (in years)", y ="Imprecision")+
  scale_x_continuous(breaks = 2:8)+
  theme_minimal()+
  theme(legend.position = "bottom")

```
```{r}
ggsave("../visuals/dev_site.png", width = 5, height = 6, bg = "white")
```

```{r}
data%>%
  group_by(targetcentralityx)%>%
  mutate(chance_upper = (targetcentralityx + 960)/2,
         chance_lower = -((targetcentralityx + 960)/2),
         center_lower = -targetcentralityx,
         center_upper = targetcentralityx)%>%
  ggplot(aes(x = targetcentralityx, y = clickdistfromtargetcenterx, col = community))+
  geom_hline(yintercept = 0, lty = 3, alpha = 0.75)+
  geom_line(aes(x = targetcentralityx, y = center_upper), lty = 1, alpha = .75, col = "firebrick")+
  geom_line(aes(x = targetcentralityx, y = center_lower), lty = 1, alpha = .75, col = "firebrick")+
  geom_line(aes(x = targetcentralityx, y = chance_upper), lty = 3, alpha = .75, col = "firebrick")+
  geom_line(aes(x = targetcentralityx, y = chance_lower), lty = 3, alpha = .75, col = "firebrick")+
  geom_point(pch = 1, alpha = 0.5)+
  scale_color_colorblind()+
  #ylim(-1500,1500)+
  theme_minimal()

```

```{r}
ptp <- data%>%
  mutate(targetposition = factor(targetposition, levels = c(1:10)))%>%
  group_by(community, targetposition)%>%
  tidyboot_mean(col= abs(clickdistfromtargetcenterx))

ggplot(ptp, aes(x = targetposition, y = mean, col = community))+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),pch = 1, alpha = 0.75, position = position_dodge(width = .5))+
  labs(x = "Binned target position", y = "Mean imprecision")+
  scale_color_colorblind()+
  #ylim(0,1000)+
  theme_minimal()
```

```{r}
pscreen1 <- data%>%
  mutate(screen = factor(screen),
         touchscreen = factor(touchscreen))%>%
  pivot_longer(cols = c(screen, touchscreen), names_to = "predictor", values_to = "value")%>%
  group_by(community, subjid, predictor, value)%>%
  summarise(mean= mean(abs(clickdistfromtargetcenterx)))

pscreen2 <- data%>%
  mutate(screen = factor(screen),
         touchscreen = factor(touchscreen))%>%
  pivot_longer(cols = c(screen, touchscreen), names_to = "predictor", values_to = "value")%>%
  group_by(community, predictor, value)%>%
  tidyboot_mean(col= abs(clickdistfromtargetcenterx))

ggplot(pscreen%>%filter(!is.na(value)), aes(x = value, y = mean, col = community))+
  geom_point(data = pscreen1%>%filter(!is.na(value)), aes(x = value, y = mean, col = community), pch = 1, alpha = .5, position = position_dodge(width = .5))+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.75, position = position_dodge(width = .5))+
  facet_grid(~predictor, scales = "free_x")+
  scale_x_discrete(labels = c("no", "yes"))+
  scale_color_colorblind()+
  #ylim(0,1000)+
  theme_few()
```

```{r}
pred <- data%>%
  group_by(community, subjid, household, children, younger_children)%>%
  summarise(mean= mean(abs(clickdistfromtargetcenterx)))%>%
  pivot_longer(cols = c(household, children, younger_children), names_to = "predictor", values_to = "value")%>%
  group_by(community,predictor)%>%
  mutate(value = scale(value))

ggplot(pred, aes(x = value, y = mean, col = community))+
  geom_point(pch = 1)+
  geom_smooth(method = "glm")+
  facet_grid(~predictor, scales = "free_x")+
  stat_cor()+
  scale_color_colorblind()+
  ylim(0,1000)+
  theme_minimal()
```