### 2022-07-20 analysis/gafo-cc-model-hier.Rmd
prep hier model run

---
title: "model"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(coda)
library(jsonlite)
library(ggthemes)
library(ggpubr)
library(data.table)
library(truncnorm)
library(matrixStats)
library(tidyboot)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

# Data

```{r}
data <- read_csv("../data/gafo-cc-clean-data.csv")
```

## Click distribution based on performance

# Model

```{r, message=FALSE, warning=F}
model <- rbindlist(
  list(
     fread("../model/output/gafo-cc-hier_gran_20_chain1.csv"),
     fread("../model/output/gafo-cc-hier_gran_20_chain2.csv")
    ), use.names=TRUE)

saveRDS(model, "../model/output/gafo-cc-model_hier.rds")
# 
# model <- readRDS("../model/output/gafo-cc-model.rds")
```

## Global Paramters

### Guessing

```{r}
model%>%
  filter(scope == "global", 
         parameter == "guessing")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_grid(community~type, scales = "free_y")+
  theme_few()+
  scale_fill_viridis_d()
```

### Bias

```{r}
model%>%
  filter(scope == "global", 
         parameter == "bias")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_grid(community~type)+
  theme_few()+
  scale_fill_viridis_d()
```

### Inference

```{r}
model%>%
  filter(scope == "global", 
         parameter == "inference")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_wrap(~type, scales = "free_y")+
  theme_minimal()+
  scale_fill_viridis_d()
```

## Community Paramters

### Guessing

```{r}
model%>%
  filter(scope == "community", 
         parameter == "guessing")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_grid(community~type, scales = "free_y")+
  theme_few()+
  scale_fill_viridis_d()
```

### Bias

```{r}
model%>%
  filter(scope == "community", 
         parameter == "bias")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_grid(community~type)+
  theme_few()+
  scale_fill_viridis_d()
```

### Inference

```{r}
model%>%
  filter(scope == "community", 
         parameter == "inference")%>%
ggplot(aes(x= value, fill = factor(chain)))+
  geom_density(alpha = .5)+
  facet_wrap(community~type, scales = "free_y")+
  theme_minimal()+
  scale_fill_viridis_d()
```

## Development

### Guessing
```{r}
guess_map <- model%>%
  filter(scope == "global", 
         parameter == "guessing")%>%
  group_by(community, type)%>%
  summarise(mode = estimate_mode(value))%>%
  pivot_wider(names_from = "type", values_from = "mode")%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(guess = plogis(intercept + slope * age))

guess_samp <- model%>%
  filter(scope == "global",
         parameter == "guessing")%>%
  pivot_wider(names_from = "type", values_from = "value")%>%
  #sample_n(1000)%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(guess = plogis(intercept + slope * age))

# guess_samp <- model%>%
#   filter(scope == "global", 
#          parameter == "guessing")%>%
#   pivot_wider(names_from = "type", values_from = "value")%>%
#   group_by(chain)%>%
#   expand_grid(age = seq(0,3, by = 0.1))%>%
#   mutate(sd_inf = exp(intercept + slope * age))%>%
#   group_by(age)%>%
#   summarise(uci = hdi_upper(sd_inf),
#             lci = hdi_lower(sd_inf))

ggplot()+
  geom_line(data = guess_samp, aes(x = age+3, y = guess, group = paste(chain,iteration)), alpha = .01)+
  #geom_line(data = inf_samp, aes(x = age+3, y = uci), lty = 2)+
  #geom_line(data = inf_samp, aes(x = age+3, y = lci), lty = 2)+
  geom_line(data = guess_map, aes(x = age+3, y = guess), col = "firebrick")+
  facet_grid(~community)+
  xlim(3,6)+
  xlab("Age")+
  ylim(0,1)+
  theme_minimal()
```

### Bias
```{r}
bias_map <- model%>%
  filter(scope == "global", 
         parameter == "bias")%>%
  group_by(community,type)%>%
  summarise(mode = estimate_mode(value))%>%
  pivot_wider(names_from = "type", values_from = "mode")%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(bias = plogis(intercept + slope * age))

bias_samp <- model%>%
  filter(scope == "global",
         parameter == "bias")%>%
  pivot_wider(names_from = "type", values_from = "value")%>%
  #sample_n(1000)%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(bias = plogis(intercept + slope * age))

# bias_samp <- model%>%
#   filter(scope == "global", 
#          parameter == "bias")%>%
#   pivot_wider(names_from = "type", values_from = "value")%>%
#   group_by(chain)%>%
#   expand_grid(age = seq(0,3, by = 0.1))%>%
#   mutate(sd_inf = exp(intercept + slope * age))%>%
#   group_by(age)%>%
#   summarise(uci = hdi_upper(sd_inf),
#             lci = hdi_lower(sd_inf))

ggplot()+
  geom_line(data = bias_samp, aes(x = age+3, y = bias, group = paste(chain,iteration)), alpha = .01)+
  #geom_line(data = inf_samp, aes(x = age+3, y = uci), lty = 2)+
  #geom_line(data = inf_samp, aes(x = age+3, y = lci), lty = 2)+
  geom_line(data = bias_map, aes(x = age+3, y = bias), col = "firebrick")+
  xlim(3,6)+
  facet_grid(~community)+
  xlab("Age")+
  ylim(0,1)+
  theme_minimal()
```

### Inference

```{r}
inf_map <- model%>%
  filter(scope == "global", 
         parameter == "inference")%>%
  group_by(community,type)%>%
  summarise(mode = estimate_mode(value))%>%
  pivot_wider(names_from = "type", values_from = "mode")%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(sd_inf = plogis(intercept + slope * age))

inf_samp <- model%>%
  filter(scope == "global",
         parameter == "inference")%>%
  pivot_wider(names_from = "type", values_from = "value")%>%
  group_by(chain)%>%
  #sample_n(1000)%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(sd_inf = plogis(intercept + slope * age))

# inf_samp <- model%>%
#   filter(scope == "global",
#          parameter == "inference")%>%
#   pivot_wider(names_from = "type", values_from = "value")%>%
#   group_by(chain)%>%
#   expand_grid(age = seq(0,3, by = 0.1))%>%
#   mutate(sd_inf = exp(intercept + slope * age))%>%
#   group_by(age)%>%
#   summarise(uci = hdi_upper(sd_inf),
#             lci = hdi_lower(sd_inf))

ggplot()+
  geom_line(data = inf_samp, aes(x = age+3, y = sd_inf, group = paste(chain,iteration)), alpha = .01, size = 0.1)+
   # geom_line(data = inf_samp, aes(x = age+3, y = uci), lty = 2)+
   # geom_line(data = inf_samp, aes(x = age+3, y = lci), lty = 2)+
  geom_line(data = inf_map, aes(x = age+3, y = sd_inf), col = "firebrick")+
  xlim(3,6)+
  facet_grid(~community)+
  labs(x = "Age", y = "Gaze inference precision (across individuals)")+
  ylim(0,1)+
  theme_minimal()
```

```{r}
ggsave("../graphs/sd_dev.png", width = 8, height = 4, scale = 1, bg = "white")
```

```{r}
alt <- model%>%
  filter(scope == "global", 
         parameter != "inference")%>%
  group_by(community,parameter,type)%>%
  summarise(mode = estimate_mode(value))%>%
  filter(type != "sigma")%>%
  pivot_wider(names_from = c("parameter", "type"), values_from = "mode")%>%
  expand_grid(age = seq(0,3, by = 0.1))%>%
  mutate(guess = plogis(guessing_intercept + guessing_slope * age),
         bias_raw = plogis(bias_intercept + bias_slope * age),
         inference = 1- guess,
         center_bias = guess * bias_raw,
         random_guessing = 1- inference - center_bias)%>%
  ungroup()%>%
  select(age,community, center_bias, random_guessing, inference)%>%
  pivot_longer(names_to = "parameter", values_to = "value", cols = c(center_bias, random_guessing, inference))%>%
  mutate(parameter = factor(parameter, levels = c("inference", "random_guessing", "center_bias")))

ggplot(alt, aes(x = age+3, y = value, fill = parameter))+
  geom_area(alpha = .9)+
  xlim(3,6)+
  xlab("Age")+
  ylim(0,1)+
  facet_grid(~community)+
  scale_fill_colorblind()+
  theme_minimal()
```

## Individual

### Guessing

```{r}
model%>%
  filter(scope == "id", 
         parameter == "guessing")%>%
  ggplot(aes(x = value, fill = factor(chain)))+
  geom_density(alpha = .25)+
  facet_wrap(~id, scales = "free")+
  theme_minimal()+
  scale_fill_colorblind()
```

### Inference

```{r}
model%>%
  filter(scope == "id", 
         parameter == "inference")%>%
  ggplot(aes(x = value, fill = factor(chain)))+
  geom_density(alpha = .25)+
  facet_wrap(~id, scales = "free")+
  theme_minimal()+
  scale_fill_colorblind()+
  xlim(0,2)
```

### Relation between paramters

```{r}
rel <- model%>%
  filter(scope == "id")%>%
  group_by(parameter, id)%>%
  summarise(mode = estimate_mode(value))%>%
  left_join(data%>%rename(id = subjID)%>%select(id, ageInYears))

ggplot(rel,aes(x = parameter, y = mode))+
  geom_line(aes(group = id), alpha = .25)+
  geom_point(col = "dodgerblue", pch = 1,alpha = .5)+
  facet_grid(~ageInYears)+
  theme_minimal()
```