### 2024-03-14 analysis/gafo-cc-model.Rmd
finalize model comparison

---
title: "model"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(coda)
library(jsonlite)
library(ggthemes)
library(ggpubr)
library(data.table)
library(truncnorm)
library(matrixStats)
library(tidyboot)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

# Data

```{r}
data <- read_csv("../data/gafo-cc-clean-data.csv")

```

# Model comparison

## Hai||om

```{r, message=FALSE, warning=F}
# haillom_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/haillom_model_comparison_chain1.csv"),
#      fread("../model/output/haillom_model_comparison_chain2.csv"),
#      fread("../model/output/haillom_model_comparison_chain3.csv"),
#      fread("../model/output/haillom_model_comparison_chain4.csv"),
#      fread("../model/output/haillom_model_comparison_chain5.csv"),
#      fread("../model/output/haillom_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "hai||om")%>%
#   saveRDS("../model/output/haillom_model_comparison.rds")

 haillom_model_comparison <- readRDS("../model/output/haillom_model_comparison.rds")
```

```{r}
haillom_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Chimfunshi

```{r, message=FALSE, warning=F}
# chimfunshi_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/chimfunshi_model_comparison_chain1.csv"),
#      fread("../model/output/chimfunshi_model_comparison_chain2.csv"),
#      fread("../model/output/chimfunshi_model_comparison_chain3.csv"),
#      fread("../model/output/chimfunshi_model_comparison_chain4.csv"),
#      fread("../model/output/chimfunshi_model_comparison_chain5.csv"),
#      fread("../model/output/chimfunshi_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "chimfunshi")%>%
#   saveRDS("../model/output/chimfunshi_model_comparison.rds")

 chimfunshi_model_comparison <- readRDS("../model/output/chimfunshi_model_comparison.rds")
```

```{r}
chimfunshi_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Mexico

```{r, message=FALSE, warning=F}
# mexico_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/mexico_model_comparison_chain1.csv"),
#      fread("../model/output/mexico_model_comparison_chain2.csv"),
#      fread("../model/output/mexico_model_comparison_chain3.csv"),
#      fread("../model/output/mexico_model_comparison_chain4.csv"),
#      fread("../model/output/mexico_model_comparison_chain5.csv"),
#      fread("../model/output/mexico_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "mexico")%>%
#   saveRDS("../model/output/mexico_model_comparison.rds")
# 
#  mexico_model_comparison <- readRDS("../model/output/mexico_model_comparison.rds")
```

```{r}
mexico_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## India

```{r, message=FALSE, warning=F}
# india_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/india_model_comparison_chain1.csv"),
#      fread("../model/output/india_model_comparison_chain2.csv"),
#      fread("../model/output/india_model_comparison_chain3.csv"),
#      fread("../model/output/india_model_comparison_chain4.csv"),
#      fread("../model/output/india_model_comparison_chain5.csv"),
#      fread("../model/output/india_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "india")%>%
#   saveRDS("../model/output/india_model_comparison.rds")

  india_model_comparison <- readRDS("../model/output/india_model_comparison.rds")
```

```{r}
india_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Akure

```{r, message=FALSE, warning=F}
# nigeria_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/akure_model_comparison_chain1.csv"),
#      fread("../model/output/akure_model_comparison_chain2.csv"),
#      fread("../model/output/akure_model_comparison_chain3.csv"),
#      fread("../model/output/akure_model_comparison_chain4.csv"),
#      fread("../model/output/akure_model_comparison_chain52.csv"),
#      fread("../model/output/akure_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "nigeria")%>%
#   saveRDS("../model/output/nigeria_model_comparison.rds")

nigeria_model_comparison <- readRDS("../model/output/nigeria_model_comparison.rds")
```

```{r}
nigeria_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Beijing

```{r, message=FALSE, warning=F}
# china_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/beijing_model_comparison_chain1.csv"),
#      fread("../model/output/beijing_model_comparison_chain2.csv"),
#      fread("../model/output/beijing_model_comparison_chain3.csv"),
#      fread("../model/output/beijing_model_comparison_chain4.csv"),
#      fread("../model/output/beijing_model_comparison_chain5.csv"),
#      fread("../model/output/beijing_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "china")%>%
#   saveRDS("../model/output/china_model_comparison.rds")

china_model_comparison <- readRDS("../model/output/china_model_comparison.rds")
```

```{r}
china_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Buenos Aires

```{r, message=FALSE, warning=F}
# argentina_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/buenos_aires_model_comparison_chain1.csv"),
#      fread("../model/output/buenos_aires_model_comparison_chain2.csv"),
#      fread("../model/output/buenos_aires_model_comparison_chain3.csv"),
#      fread("../model/output/buenos_aires_model_comparison_chain4.csv"),
#      fread("../model/output/buenos_aires_model_comparison_chain5.csv"),
#      fread("../model/output/buenos_aires_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "argentina")%>%
#   saveRDS("../model/output/argentina_model_comparison.rds")

argentina_model_comparison <- readRDS("../model/output/argentina_model_comparison.rds")
```

```{r}
argentina_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Leipzig

```{r, message=FALSE, warning=F}
# leipzig_model_comparison <- rbindlist(
#   list(
#      fread("../model/output/leipzig_model_comparison_chain1.csv"),
#      fread("../model/output/leipzig_model_comparison_chain2.csv"),
#      fread("../model/output/leipzig_model_comparison_chain3.csv"),
#      fread("../model/output/leipzig_model_comparison_chain4.csv"),
#      fread("../model/output/leipzig_model_comparison_chain5.csv"),
#      fread("../model/output/leipzig_model_comparison_chain6.csv")
#     ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "leipzig")%>%
#   saveRDS("../model/output/leipzig_model_comparison.rds")

leipzig_model_comparison <- readRDS("../model/output/leipzig_model_comparison.rds")
```

```{r}
leipzig_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## Auckland

```{r, message=FALSE, warning=F}
# auckland_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/auckland_model_comparison_chain1.csv"),
#     fread("../model/output/auckland_model_comparison_chain2.csv"),
#     fread("../model/output/auckland_model_comparison_chain3.csv"),
#     fread("../model/output/auckland_model_comparison_chain4.csv"),
#     fread("../model/output/auckland_model_comparison_chain5.csv"),
#     fread("../model/output/auckland_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "auckland")%>%
#   saveRDS("../model/output/auckland_model_comparison.rds")

auckland_model_comparison <- readRDS("../model/output/auckland_model_comparison.rds")
```

```{r}
auckland_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## bandongo

```{r, message=FALSE, warning=F}
# bandongo_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/bandongo_model_comparison_chain1.csv"),
#     fread("../model/output/bandongo_model_comparison_chain2.csv"),
#     fread("../model/output/bandongo_model_comparison_chain3.csv"),
#     fread("../model/output/bandongo_model_comparison_chain4.csv"),
#     fread("../model/output/bandongo_model_comparison_chain5.csv"),
#     fread("../model/output/bandongo_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "bandongo")%>%
#   saveRDS("../model/output/bandongo_model_comparison.rds")
# 
bandongo_model_comparison <- readRDS("../model/output/bandongo_model_comparison.rds")
```

```{r}
bandongo_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## bayaka

```{r, message=FALSE, warning=F}
# bayaka_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/bayaka_model_comparison_chain1.csv"),
#     fread("../model/output/bayaka_model_comparison_chain2.csv"),
#     fread("../model/output/bayaka_model_comparison_chain3.csv"),
#     fread("../model/output/bayaka_model_comparison_chain4.csv"),
#     fread("../model/output/bayaka_model_comparison_chain5.csv"),
#     fread("../model/output/bayaka_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "bayaka")%>%
#   saveRDS("../model/output/bayaka_model_comparison.rds")

bayaka_model_comparison <- readRDS("../model/output/bayaka_model_comparison.rds")
```

```{r}
bayaka_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## khwe

```{r, message=FALSE, warning=F}
# khwe_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/khwe_model_comparison_chain1.csv"),
#     fread("../model/output/khwe_model_comparison_chain2.csv"),
#     fread("../model/output/khwe_model_comparison_chain3.csv"),
#     fread("../model/output/khwe_model_comparison_chain4.csv"),
#     fread("../model/output/khwe_model_comparison_chain5.csv"),
#     fread("../model/output/khwe_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "khwe")%>%
#   saveRDS("../model/output/khwe_model_comparison.rds")

khwe_model_comparison <- readRDS("../model/output/khwe_model_comparison.rds")
```

```{r}
khwe_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## plymouth

```{r, message=FALSE, warning=F}
# plymouth_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/plymouth_model_comparison_chain1.csv"),
#     fread("../model/output/plymouth_model_comparison_chain2.csv"),
#     fread("../model/output/plymouth_model_comparison_chain3.csv"),
#     fread("../model/output/plymouth_model_comparison_chain4.csv"),
#     fread("../model/output/plymouth_model_comparison_chain5.csv"),
#     fread("../model/output/plymouth_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "plymouth")%>%
#   saveRDS("../model/output/plymouth_model_comparison.rds")

plymouth_model_comparison <- readRDS("../model/output/plymouth_model_comparison.rds")
```

```{r}
plymouth_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## stanford

```{r, message=FALSE, warning=F}
# stanford_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/stanford_model_comparison_chain1.csv"),
#     fread("../model/output/stanford_model_comparison_chain2.csv"),
#     fread("../model/output/stanford_model_comparison_chain3.csv"),
#     fread("../model/output/stanford_model_comparison_chain4.csv"),
#     fread("../model/output/stanford_model_comparison_chain5.csv"),
#     fread("../model/output/stanford_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "stanford")%>%
#   saveRDS("../model/output/stanford_model_comparison.rds")

stanford_model_comparison <- readRDS("../model/output/stanford_model_comparison.rds")
```

```{r}
stanford_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## turkey

```{r, message=FALSE, warning=F}
# turkey_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/turkey_model_comparison_chain1.csv"),
#     fread("../model/output/turkey_model_comparison_chain2.csv"),
#     fread("../model/output/turkey_model_comparison_chain3.csv"),
#     fread("../model/output/turkey_model_comparison_chain4.csv"),
#     fread("../model/output/turkey_model_comparison_chain5.csv"),
#     fread("../model/output/turkey_model_comparison_chain6.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "turkey")%>%
#   saveRDS("../model/output/turkey_model_comparison.rds")

turkey_model_comparison <- readRDS("../model/output/turkey_model_comparison.rds")
```

```{r}
turkey_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## uganda

```{r, message=FALSE, warning=F}
# uganda_model_comparison <- rbindlist(
#   list(
#     fread("../model/output/uganda_model_comparison_chain1.csv"),
#     fread("../model/output/uganda_model_comparison_chain2.csv"),
#     fread("../model/output/uganda_model_comparison_chain31.csv"),
#     fread("../model/output/uganda_model_comparison_chain4.csv"),
#     fread("../model/output/uganda_model_comparison_chain5.csv"),
#     fread("../model/output/uganda_model_comparison_chain61.csv")
#   ), use.names=TRUE)%>%
#   group_by(chain, iteration, model)%>%
#   summarise(like = sum(loglike))%>%
#   mutate(community = "uganda")%>%
#   saveRDS("../model/output/uganda_model_comparison.rds")

uganda_model_comparison <- readRDS("../model/output/uganda_model_comparison.rds")
```

```{r}
uganda_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

## windhoek

```{r, message=FALSE, warning=F}
# windhoek_model_comparison <- rbindlist(
# list(
#   fread("../model/output/windhoek_model_comparison_chain1.csv"),
#   fread("../model/output/windhoek_model_comparison_chain2.csv"),
#   fread("../model/output/windhoek_model_comparison_chain3.csv"),
#   fread("../model/output/windhoek_model_comparison_chain4.csv"),
#   fread("../model/output/windhoek_model_comparison_chain5.csv"),
#   fread("../model/output/windhoek_model_comparison_chain6.csv")
# ), use.names=TRUE)%>%
# group_by(chain, iteration, model)%>%
# summarise(like = sum(loglike))%>%
# mutate(community = "windhoek")%>%
# saveRDS("../model/output/windhoek_model_comparison.rds")

windhoek_model_comparison <- readRDS("../model/output/windhoek_model_comparison.rds")
```

```{r}
windhoek_model_comparison%>%
  group_by(community, model,chain)%>%
  summarise(margllh = logSumExp(like), 
            n = n_distinct(iteration))%>%
  pivot_wider(names_from = model, values_from = margllh)
```

# Summary

```{r}
bf_comparison <- bind_rows(
  haillom_model_comparison,
  chimfunshi_model_comparison,
  mexico_model_comparison,
  india_model_comparison,
  nigeria_model_comparison,
  china_model_comparison,
  argentina_model_comparison,
  leipzig_model_comparison,
  auckland_model_comparison,
  bandongo_model_comparison,
  bayaka_model_comparison,
  khwe_model_comparison,
  plymouth_model_comparison,
  stanford_model_comparison,
  turkey_model_comparison,
  windhoek_model_comparison,
  uganda_model_comparison
) %>%
  group_by(community, model)%>%
  summarise(margllh = logSumExp(like))%>%
  pivot_wider(names_from = model, values_from = margllh)%>%
  summarise(bf_inf_center = inf - center, 
            bf_inf_guess = inf - guess)

saveRDS(bf_comparison, "../model/output/bf_comparison.rds")

```