### 2024-03-14 paper/supplement.Rmd

---
title: "Unity in diversity: Children from 17 diverse communities process gaze in similar ways"
subtitle: "Supplemental material"
author: "..."
output: 
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
bibliography: library.bib
#csl: pnas.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\figurename}{Supplementary Figure}
  \renewcommand{\tablename}{Supplementary Table}
---

```{r, echo=FALSE, message=F}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(ggpubr)
library(tidyboot)
library(tidybayes)
library(maps)
library(brms)
library(cowplot)
library(ggridges)
library(papaja)
library(coda)
library(ggrepel)
library(knitr)
library(kableExtra)
library(pals)
library(RColorBrewer)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, size="small")
options(knitr.kable.NA = '')

set.seed(1337)
```

```{r}
data <- read_csv("../data/gafo-cc-clean-data.csv")%>%
  mutate(continent = ifelse(
    community %in% c("leipzig", "plymouth"), "Europe",
    ifelse(
      community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda"), "Africa",
      ifelse(community %in% c("stanford", "mexico", "buenos_aires"), "Americas",
             ifelse(community %in% c("india", "beijing", "turkey"), "Asia", "Oceania"))
    )
  ))%>%
  mutate(continent = factor(continent, levels = c("Americas", "Africa", "Europe", "Asia", "Oceania"), ordered = T))%>%
  mutate(community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)",
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)",
                            uganda = "Nyabyeya (Uganda)"))
```
```{r}
qual_col_pals = brewer.pal.info[brewer.pal.info$colorblind == TRUE,]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
colors=kelly(n=n_distinct(data$community)+1)
colors <- colors[ !colors == '#F2F3F4']
```

# Overview

This docuemnt provides a detailed overview of the participants and methods of the in the section [site descriptions](#site-descriptions)

# Participants

```{r demtable, echo=F}
demtab <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent)

kableExtra::kable(demtab,
             col.names = c("Community", "Country", "Continent", "N", "Sex (male)", "Age (range)"),
              caption = "Participant demographics.", 
             format="latex", booktabs=TRUE
             )
```

```{r}
# page <- data%>%
#   group_by(community,continent)%>%
#   distinct(subjid, .keep_all = T)%>%
#   mutate(dens = approxfun(density(ageinyears))(ageinyears))%>%
#   ggplot(aes(y = community, x = ageinyears))+
#   #facet_grid(continent~., scales = "free", space = "free_y")+
#   geom_density_ridges(
#     jittered_points = TRUE,
#     position = position_points_jitter(width = 0, height = 0),
#     point_shape = '|',  point_alpha = 0.5, alpha = 0.7,
#   )+
#   labs(x = "Age (in years)", y= "")+
#   theme_few()+
#   theme(strip.background = element_blank(),strip.text.y = element_blank())
# 
# pred_agg_i <- data%>%
#   mutate(older_children = children - 1 - younger_children)%>%
#   group_by(community, subjid)%>%
#   summarise(screen = mean(screen, na.rm = T),
#             touchscreen = mean(touchscreen, na.rm = T),
#             older_children = mean(older_children, na.rm = T),
#             household = mean(household, na.rm = T))
# 
# pred_agg_g <- data%>%
#   mutate(older_children = children - 1 - younger_children)%>%
#   group_by(community, continent)%>%
#   summarise(screen = mean(screen, na.rm = T),
#             touchscreen = mean(touchscreen, na.rm = T),
#             older_children = mean(older_children, na.rm = T),
#             younger_children = mean(younger_children, na.rm = T),
#             children = mean(children, na.rm = T),
#             household = mean(household, na.rm = T))%>%
#   pivot_longer(names_to = "dem", values_to = "value", cols = c(screen, touchscreen, older_children,children, younger_children, household))
# 
# ppred <- pred_agg_g%>%
#   filter(dem == "screen" | dem == "touchscreen")%>%
#   ggplot(aes(y = community, x = dem, fill = value))+
#   geom_point(pch = 22, size = 10, stroke = 0)+
#   #facet_grid(continent~., scales = "free", space = "free_y")+
#   labs(x = "", y= "")+
#   scale_fill_viridis_c(name = "Prevalence")+
#   theme_few()+
#   scale_x_discrete(guide = guide_axis(n.dodge = 2))+
#   theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), strip.text.y.right = element_text(angle = 0))
```

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(data$country)` different countries. Table \@ref(tab:demtable) gives the sample size per community together with some basic demographic information. Figure \@ref(fig:figpred1)A shows the age distribution in each setting. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category. Additional children were recruited but did not contribute data to the study. The recruitment strategy varied from community to community and we therefore cannot provide detailed information about the number of children that were excluded. All children who contributed at least four valid data points were included in the final sample, however, we excluded children when parents did not fill out the questionnaire correctly (e.g., when the number of children living in the household was reported to be larger than the household itself or when the number of children reported to live in the household equaled the number of children younger than the child being tested). Twelve children had to be excluded for one of these reasons (one from Akure, two from Stanford, seven from Pune and two from Windhoek). We did not exclude any participants for performance reasons. A detailed description of each site and the way children were recruited can be found below.

## Site descriptions

The descriptions below give detailed information about each data collection site, including the strategy used for recruiting children. The order is alphabetically.

### Akure (Nigeria) - Location

Akure is situated in the Ondo State in southwest Nigeria. About 800,000 people live in this medium-sized city, and most of them are from the nearby urban and rural areas.

There are two different seasons in Akure's tropical climate: the wet season, which runs from April to October, and the dry season, which runs from November to March. Rich vegetation, including deep forests and productive agriculture, is what defines the city.

Akure's natural surroundings are enhanced by the abundant vegetation that includes dense forests, beautiful farmlands, and vivid flora, which is fostered by the climate and fertile soils.

Akure boasts a diverse populace, representing a cultural mosaic where the Yoruba, Igbo, and Hausa communities coexist harmoniously. While the Yoruba ethnicity predominates, English serves as a widely spoken language alongside the native tongue.

Over centuries, Akure has cultivated a sense of unity among its residents, transcending ethnic boundaries and nurturing a strong community spirit grounded in mutual respect and cooperation. This communal ethos extends from familial bonds – encompassing both nuclear and extended families – to broader social networks within neighborhoods, workplaces, and religious congregations. Akure's evolution from a traditional Yoruba kingdom to a thriving urban metropolis is deeply embedded in history and has a rich cultural legacy. The landscape is dotted with evidence of old kingdoms and communities, highlighting the city's lasting influence and crucial role in the history of the area.

The primary industry of Akure is agriculture, with a large number of residents working in a variety of farming-related fields including agricultural cultivation, animal husbandry, and the extraction of palm oil. In addition to providing for livelihoods, this traditional subsistence approach maintains cultural customs and connections to the land, tying communities to their agricultural origins.

The variety of architectural styles and types of homes found in the city are reflected in the housing landscape. Akure offers a diverse range of housing options, including modern urban dwellings scattered across the city and ancient mud-brick homes tucked away in rural hamlets. These housing options can accommodate nuclear families, extended kin networks, and communal living arrangements.

A set of fundamental principles that are deeply embedded in the fabric of Akure society define parenting there. Parents are essential in passing on cultural legacy and forming the identities of future generations because they emphasize reverence for elders, community cohesion, and the value of education. Cultural festivals, ceremonies, and rite of passage serve as platforms for imparting values and fostering social cohesion within the community.

The differences in parents' educational attainment in Akure are a reflection of their socioeconomic status and differences in access to formal education. Childcare practices often involve extended family members, particularly grandparents, who play significant roles in nurturing and guiding children.
The city is home to a wide range of educational institutions including public and private schools, that provide extensive curriculum that incorporate standard academic disciplines along with enrichment and cultural education initiatives.

Akure has embraced technology in recent years, with a notable increase in digital adoption, especially among its younger population. The widespread availability of reasonably priced smartphones and other touchscreen devices has contributed to the growing integration of technology into everyday life, affecting both urban and peri-urban areas. This technological revolution has changed social interactions and cultural dynamics within the community by democratizing access to information, communication, and entertainment.

For this study, participants were recruited from churches during Sunday services and residences during community visits within the city of Akure, ensuring a wide representation of its vibrant population.

**References and further readings:**

* ttps://worldpopulationreview.com/world-cities/akure-population

### Auckland | Tāmaki Makaurau (New Zealand) - Location

Auckland | Tāmaki Makaurau is the largest city of New Zealand | Aotearoa with a population of about 1.5 million people. The city is located on the North Island | Te Ika-a-Māui inbetween the Hauraki Gulf | Tīkapa Moana to the East and the Waitākere Ranges to the West. The climate is defined as either oceanic (Cfb) or subtropical (NIWA). Temperatures are mild throughout the year, with an average temperature of 23.7 during summer and 14.7 during winter months.

The median age in the Auckland region was 34.7 years. 181,194 Māori people with a median age of 24.9 years were counted in the 2018 census. Overall the population is composed of 53.5% European, 28.2% Asian, 15.5% Pacific Peoples, 11.5% Māori, 2.3% Middle Eastern / Latin American / African and 1.1% Other ethnicity. 92.7% are English speakers, with 4.4% speaking Northern Chinese (Mandarin Chinese), 4.4% speaking Samoan and 2.4% speaking te reo Māori.

Māori peoples arrived in Auckland | Tāmaki Makaurau around 1350 from other Pacific Islands. Before the arrival of European New Zealanders | Pākehā New Zealand was populated by about 20,000 Māori. Aotearoa was colonized by the Dutch and British Empires in the 18th Century. The effects the British Monarchy had on Aotearoa have been at the center of long lasting societal disputes. The Treaty of Waitangi | Te Tiriti o Waitangi is a document signed by 240 Māori chiefs and representatives of the British Crown in 1840 trying to establish a cooperative relationship while redeeming some of the mistakes that have been made. This document is still important for the revitalization of Māori culture throughout Aotearoa.

51.9% of the people in Auckland are employed full time, with an unemployment rate of 4.1% and 30.4% not in the labor force (i.e. retirees). The main occupations are professionals (25.9%) and managers (18.1%) with only 5.9% operating machinery and 7.9% laborers. The people of Auckland are part of a service society with a median income of 34,400NZD and 19.8% of the population earning more than 70,000NZD.

Auckland is made up of mainly sub-urban areas with single family housing. 45.4% of the population own or partly own a house. People mostly live with their core family in loosely separated areas of the city. Those areas commonly have an economical center with shops and restaurants as well as a mix of private and public education institutions.  The socialization goals and cultural norms are comparable to most western, democratic societies. High education standards are deemed as vital for success. Success is determined by career paths, income and general success in professional and private aspects of life. Sport plays a central role in most aspects of society. The role of sports is viewed as essential for living a healthy and successful life and as such is a central part in education programs.

Education levels are generally good in Auckland. 18.6% of the population hold a Bachelor's Degree. Interestingly about 9.1% move overseas for their secondary education. Parenting and childcare practises are comparable to other western cultures, with some adjustments being implemented in recent years to include more Māori cultural practises in education programs. Such as starting the day or specific activities | hui with Karakia, a form of incantation to ensure a favorable outcome to important events and undertakings. (see Literature for further information on early childhood education in New Zealand.)

Aotearoa is an industrialized nation, nearly everybody has a smartphone and uses technology in their everyday life. Many administrative processes are very much digitized. There is no perceived mistrust towards technology and new developments are generally embraced with great interest.

For this study we reached out to early childcare centers and schools. Two early childcare centers and one primary school agree to participate. Some additional children were tested during home visit sessions. All three educational institutions are visited by children from different cultural and socio-economic backgrounds. One early childcare center is affiliated with the University of Auckland and prioritizes children from University students and staff in their admission process. Testing sessions took place in separated spaces (office or break rooms) with one participant at a time. Participants received a certificate “for becoming a junior scientist” and a small gift (coloring book or sticker set) after their participation. (All numbers in the above description are from the 2018 census data report.)

**References and further readings:**

* Gunn, A. C., Child, C., Madden, B., Purdue, K., Surtees, N., Thurlow, B., & Todd, P. (2004). Building inclusive communities in early childhood education: Diverse perspectives from Aotearoa/New Zealand. Contemporary Issues in Early Childhood, 5(3), 293-308.
* Shuker, M. J., & Cherrington, S. (2016). Diversity in New Zealand early childhood education: Challenges and opportunities. International Journal of Early Years Education, 24(2), 172-187.
* Tesar, M. (2015). New Zealand perspectives on early childhood education: Nāku te rourou nāu te rourou ka ora ai te iwi. Journal of Pedagogy, 6(2), 9-18.
* New Zealand Census 2018 for Auckland Region: https://www.stats.govt.nz/tools/2018-census-place-summaries/auckland-region#population-and-dwellings

```{r figpred1, include = T, fig.align = "center", fig.cap = "A) Age distribution and B) prevalence of different types of technology by community. Left: dots show individual data points. Right: color shows the proportion of children with access to screens or touchscreens.", out.width="100%"}
# techpred <- plot_grid(page, ppred, rel_widths = c(1, 0.5),align = "h", labels = c("A", "B"))
# 
# ggsave(plot = techpred, "../visuals/techpred.png",  height = 7, width = 10, scale = 1)

knitr::include_graphics("../visuals/techpred.png")

```

### Bandongo (Rep. Congo) - Ethnic group

**References and further readings:**

* ...

### Bayaka (Rep. Congo) - Ethnic group

**References and further readings:**

* ...

### Beijing (China) - Location

Beijing, a municipality directly under the Central Government, is the capital of the People's Republic of China and the country's political, economic, cultural and transportation center as well as a famous historic city and a modern international city. Beijing covers an area of 16,410.54 square kilometers (sq.km) with a population of 21,843,000 in 2022. Beijing ranked second in Chinese cities in terms of GDP in 2022. Geographically, Beijing is located in the northern part of the North China Plain.

Beijing is an inland city and is surrounded by mountains to its west, north and northeast. Mountain areas occupy about 62% of the municipality's total area. The land is characterized by urban and suburban areas, with parks, squares, and historical buildings located mostly in the downtown districts, and flat agricultural land and mountains in the suburban districts. Beijing has a temperate, continental monsoonal climate, characterized by short spring and autumns, hot and rainy summers, and cold and dry winters. Its average annual temperature is 11 to 14oC. The coldest month in Beijing is January, with an average temperature of -9 to -4°C; the hottest month is July, with an average of 25°C. Around 75 percent of its annual precipitation occurs in summer, with the heaviest rainfall in July and August.

According to the 2021 National Census data, in Beijing, there were 20845166 (95%) Han Chinese along with 1047929 (4.8%) people in other minority nationalities (e.g., Manchu, Hui ethnic, Mongolian ethnic). Mandarin Chinese is commonly spoken by all ethnic groups though many of the minority nationalities also have their own national languages, such as Mongolian language used by Mongolian ethnic. (4)

Beijing is one of the birthplaces of Chinese Civilization and one of the five ancient capital cities in China. With but few interruptions, Beijing has been the capital of China for some eight centuries: including the capital of the State of Yan during the Western Zhou Dynasty (11th Century-771BC), the capital of Jin Dynasty (1153-1215), the capital of Yuan Dynasty (1271-1368), the capital of Ming Dynasty (1421-1644), and the capital of Qing Dynasty (1644-1912). In 1949, Beiping (former name of Beijing) was renamed Beijing and became the capital of the People’s Republic of China. Since then, it gradually develops into a modern metropolis.

From the early 1950s to the reform and opening up, Beijing aimed to build a modernized industrial base, and the secondary industry became the dominant industry in economic development. In 1954, the proportion of the secondary industry was 44.1%, exceeding the tertiary industry for the first time, and in 1978, the proportion reached 71%. Since the 1980s, Beijing has vigorously developed the tertiary industry, and the proportion of the tertiary industry once again exceeded that of the secondary industry in 1994, and the proportion of the tertiary industry exceeded 60% in 1998, 70% in 2005, and 80% in 2016, reaching 84% in 2022, making it the city's economic "leader". According to the statistical communiqué on the National Economy and Social Development of Beijing in 2022 the top three sectors in terms of absolute volume of GDP are finance (19.7%), information transmission, software and information technology services (17.9%) and industry (12.1%).

Beijing has various types of residential areas, ranging from traditional courtyard houses (siheyuan) in hutongs (narrow alleyways) to modern apartment buildings, gated residential complexes and detached houses. Household composition varies, ranging from single individuals and couples to families with children and multi-generational households. Notably, it is a strong tradition for grandparents to be involved in grandchildren care and household tasks. In big cities, such as Beijing, the involvement of grandparents in grandchild care is especially helpful since it relieves the high-pressure of a two-career family. Since China fully implemented the “two-child-per-family” policy in 2016, there are now usually one or two children per family in Beijing.

Parental socialization goals in Beijing (as well as in other cities) align with broader Chinese Confucianism cultural norms and typically emphasize the following seven developmental goals for children: Gaining knowledge, internalizing social norms, being modest about success and ashamed of failure, displaying self-restraint and filial piety, and maintaining harmonious relationships with others.  Education and learning are seen as very important aspects of children’s lives. For primary school children in Beijing, parents (also the education system) especially emphasize on academic achievement. Moreover, with China’s growing exposure to western values, autonomy has been promoted by current educationalists during the last decade, and is especially emphasized in developed cities such as Beijing. As to parental education levels, according to the data of 2020 population census, the average years of schooling for permanent population in Beijing aged 15 and above is 12.64 years, and among whom 48% have university degree (15 years of education and above). So, the percentage for the parents holding a university degree should be much higher than 48%.

It’s common that parents spend large parts of the day away from home and children are cared for externally. Between 2 and 3 years of age young children can attend nurseries (very few nurseries accept children as young as 1.5-year-old). Nurseries are largely commercialized and non-subsidized. Starting at age 3, children can attend either publicly subsidized kindergarten or private kindergarten. Children move on to primary school at six. Public primary education is free and compulsory. There are also private primary schools that cater to other groups, such as international primary schools that serve families catering to the expatriate community and families seeking an international education, as well as those who want to get prepared for studying abroad later. The samples of the current study were drawn from public kindergarten and public primary school.

Children growing up in Beijing have a high exposure to technology. Smartphones and other touchscreen devices are prevalent in most families in Beijing and are used widely for education, communication, and entertainment. In 2022, almost every Chinese citizen accessed the internet via mobile devices and Beijing has been fully covered by ultra fast 5G signals.

95% of the sampled Children were recruited from local public kindergarten and primary school. We firstly introduced the experiment to the heads of the kindergarten/school. Once getting their permission, we asked the teachers to help send the printed letter (information about the study, informed consent, as well as the five questions) to the parents. And a Mandarin speaking research assistant then tested kids whose parents signed the informed consent in a separate room in the kindergarten/school. A few of the youngest kids (5% of total) were recruited by directly introducing the experiment to the parents who we personally knew and tested the kids either in a separate room at our institute or at their home.

**References and further readings:**

* https://english.beijing.gov.cn/beijinginfo/index.html
* https://www.chinadiscovery.com/beijing-tours/maps/beijing-districts-map.html
* https://www.britannica.com/place/Beijing/Landscape; http://english.igsnrr.cas.cn/eocg/moc/200704/t20070409_8626.html
* http://us.mofcom.gov.cn/article/aboutchina/202011/20201103012649.shtml
* http://tjj.beijing.gov.cn/EnglishSite/
* Binah-Pollak, A. (2014). Discourses and practices of child-rearing in China: The bio-power of parenting in Beijing. China Information, 28(1), 27-45. doi:10.1177/0920203x13517617
* Chen, F., & Liu, G. (2011). The Health Implications of Grandparents Caring for Grandchildren in China. The Journals of Gerontology: Series B, 67B(1), 99-112. doi:10.1093/geronb/gbr132
* Chen, F., Short, S. E., & Entwisle, B. (2000). The impact of grandparental proximity on maternal childcare in China. Population Research and Policy Review, 19(6), 571-590. doi:10.1023/A:1010618302144
* Luo, R., Tamis-LeMonda, C. S., & Song, L. (2013). Chinese parents’ goals and practices in early childhood. Early Childhood Research Quarterly, 28(4), 843-857. doi:https://doi.org/10.1016/j.ecresq.2013.08.001
* Nyland, B., Zeng, X., Nyland, C., & Tran, L. (2009). grandparents as educators and carers in China. Journal of Early Childhood Research, 7(1), 46-57. doi:10.1177/1476718x08098353
* Xu, H. (2019). Physical and mental health of Chinese grandparents caring for grandchildren and great-grandparents. Social Science & Medicine, 229, 106-116. doi:https://doi.org/10.1016/j.socscimed.2018.05.047
* Zhu, N., Chang, L. (2019). Education and Parenting in China. In: Sorbring, E., Lansford, J. (eds) School Systems, Parent Behavior, and Academic Achievement. Young People and Learning Processes in School and Everyday Life, vol 3. Springer, Cham. https://doi.org/10.1007/978-3-030-28277-6_2

### Buenos Aires (Argentina) - Location

Buenos Aires is the capital of Argentina and located in the North-East of the country, on the shores of the Rio de la Plata. Administratively, the city is autonomous (Autonomous City of Buenos Aires) and has a separate administration from the surrounding Province of Buenos Aires. The city has a population of about 3 million inhabitants and the larger metropolitan area of about 15 million inhabitants. The majority of the population are either of European descent or of mixed descent, and about 2% of the population is of indigenous descent (e.g., Quechua, Guaraní, Mapuche, Aymara). Spanish is the predominant language in Buenos Aires, and the local dialect is known as Rioplatense Spanish.

The city has a humid, subtropical climate and has generally mild weather. The average temperatures range from 11 degrees Celsius (July) to 25 degrees (January), but heat waves are common during the summer months (December to February).

Buenos Aires was first established during Spanish colonization in the 16th century and was involved in trans-Atlantic trade when it became the capital of the Viceroyalty of Rio de la Plata (1776). The War of Independence against the Spanish crown lasted from 1810 to 1816 and was followed by an internal civil war between federalists and centralists over the constitution of the nation state. In the present day, Argentina is a presidential federal republic and Buenos Aires is its political and cultural center.

Moreover, the city is the commercial, financial, and industrial hub of Argentina and a major site for tourism. Buenos Aires has a market economy and the main economic activities center on the service sector (e.g. finance, commerce, tourism, social services) and, to some extent, manufacturing. The city contributes about 20% to the Gross Domestic Product of Argentina. Most people in Buenos Aires live in apartment buildings, but informal settlements have been growing over the last decades. Argentina has experienced one of the highest inflation rates in the world in the past years and poverty has been steadily increasing—though poverty rates in the city of Buenos Aires tend to be lower than in the surrounding Buenos Aires Province.

Commonly, children grow up in nuclear families, but ties with extended family tend to be strong. Generally, family and closeness to family are important cultural values in Argentina. Traditionally, female caregivers had a central role in the family, but this has been undergoing change in recent years, particularly, in middle-class families, with male caregivers taking on caring responsibilities, too. Argentina was the first country in South America to legalize same-sex marriage in 2010 and Buenos Aires in particular is perceived as progressive and liberal city. Some parenting values in middle-class families in Argentina, such as being responsive to the child and providing stimulating encounters, are similar to countries such as the U.S. or France. Caregivers tend to control their children’s schedules/activities, but children do not necessarily perceive this control as negative and caregivers are more permissive with younger children.

Caregivers in the current sample mostly came from a middle-class background and usually had completed tertiary education. In Argentina, children attend nursery/kindergarten from two to three years of age with the last year of kindergarten, from age 5, being compulsory. Children attend primary school from age 6 years and secondary school from age 12 (secondary school lasts 5 or 6 years). Children are schooled in Spanish and usually learn a second language (usually English) from the age of 9. Children in Indigenous communities in other Argentinian Provinces (e.g. Salta Province, Formosa Province) will often speak their respective Indigenous language (e.g. Wichí) as their primary language and schools often rely on bilingual teaching assistants to help during schools lessons (that are predominantly taught in Spanish).
Smartphones, tablets, and computers are common, and children are exposed to these devices from early on.
Children were recruited from one local kindergarten by sending out information letters and consent forms to parents. Children who had parental consent to take part in the study were tested in Spanish in a quiet room in their kindergarten by a local experimenter.

**References and further readings:**

* de Minzi, M. C. R., Lemos, V., Rubilar, J. V. (2014). Argentine culture and parenting styles in Parenting Across Cultures: Childrearing, Motherhood and Fatherhood in Non-Western Cultures, Ed. Selin, H., Springer, 277–292.

* Bornstein, M. H. (1996). Ideas about parenting in Argentina, France, and the United States. International Journal of Behavioural Development, 19, 347–368.

* Buenos Aires Cuidad. Useful information on Buenos Aires. Retrieved from: https://buenosaires.gob.ar/solarcities/en/conoce-buenos-aires

* Benitez, J. (2022). Moving on up? The downside of the high life in Buenos Aires social housing. The Sociological Review. Retrieved from: https://thesociologicalreview.org/magazine/october-2022/verticality/moving-on-up/

* Giménez, J. (2023). La pobreza aumentó al 40.1% en el primer semestre de 2023 y afecta al 56.2% de los niños. Retrieved from: https://chequeado.com/el-explicador/la-pobreza-aumento-al-401-en-el-primer-semestre-de-2023-y-afecta-al-562-de-los-ninos/

* Instituto Nacional de Estadística y Censos, Censo Nacional de Población, Hogares y Viviendas (2022). Retrieved from: https://www.indec.gob.ar/indec/web/Nivel4-Tema-2-41-135

### Chimfunshi (Sambia) - Location

Chimfunshi wildlife orphanage is an NGO providing a home for about 150 chimpanzees. The sanctuary is located in the Copperbelt region in the northern part of Zambia close to the DRC border. The chimpanzees are housed in eight different enclosures; each of an area up to 121.5Ha. The Sanctuary provides education, housing and medical care for the employees and their families which are in total about 300 people living in 5 different compounds around Chimfunshi.

Chimfunshi itself was founded in 1983 by David and Sheila Siddle who lived in Zambia since the 1950’s. In the past years they rehabilitated chimpanzees who were confiscated from poachers and from obsolete zoos and circuses worldwide. To ensure long-term sustainability, the sanctuary is managed by a Board of Trustees.

Zambia experiences a predominantly subtropical climate. The temperatures throughout the year are relatively mild but can be cold in winter due to the altitude of the country in the north. The countries’ climate oscillates between a cool dry season (May to August) followed by a short hot and dry season which turns into a hot and humid rainy season lasting from November to April. The sanctuary is nestled amidst woodland and savanna, close to the Kafue River.

The country, with approximately 18 million inhabitants, consists of a diverse population primarily composed of Bantu-speaking groups. Among these, the Bemba ethnicity stands out as the largest, accompanied by other significant groups such as Chewa, Tonga, and Lozi. While English serves as the official language, it is spoken natively by only around 2% of the population. However, for those with higher education levels, English becomes the predominant second language. The patterns of migrations has contributed to a rich tapestry of linguistic and cultural diversity across the nation., Due to a robust mining industry in the Copperbelt region, the population hailing from various corners of Zambia and even neighboring countries.

Zambia, predominantly a Christian nation, encompasses a diverse religious landscape. Due to the presence of an Asian community, Hindu and Muslim are represented beside the Christian denominations including Protestant, Roman Catholic, Jehovah’s Witnesses, and various fundamentalist churches. Cultural traditions from the Congo basin significantly influence the northeast and northwest regions of Zambia. Despite modernization, Zambian society continues to uphold traditional norms within families, villages, and clans. These enduring values emphasize mutual respect and reciprocity as cornerstones of communal harmony.

Agriculture and copper mining and smelting, particularly in the northern Copperbelt region, are the primary sectors of Zambia’s economy. Despite the abundance of copper deposits, Zambia ranks among the world’s poorest nations. The majority of the impoverished, three-fourths to be precise, reside in rural areas. Similar to other African nations, there’s a pronounced trend of migration from rural to urban areas, with men migrating more frequently while women often stay back in the villages. The architectural style in rural areas significantly differs across ethnic groups, with construction materials ranging from mud and thatch to brick and other resources.

People employed at the Chimfunshi wildlife orphanage reside in compounds. These compounds are small communities comprised of stone houses topped with corrugated iron roofs. They feature wells for clean water access, solar systems for power, and frequently, small gardens for crop cultivation and/or livestock rearing, such as chickens.

Since its Independence Day, Zambia has made significant investments in education. Currently, education is mandatory for children aged 7 to 14. The educational system is structured with a 7-year primary school program followed by a 5-year secondary level. In 2011, the Twampane School was established at Chimfunshi, providing the children of its employees the chance to pursue education up to the 9th grade. Approximately 120 students reside within the Chimfunshi area, while the rest commute from nearby farms and settlements, bringing the total student population to around 220. Sponsorship programs are in place to support academically outstanding students in attending secondary school. Additionally, Chimfunshi provides on-site vocational training and further education for students in the final grades.

The availability of touchscreen electronic devices like smartphones or laptops depends on each family’s financial capacity. While many still utilize cellphones, the lack of smartphone ownership can pose challenges in accessing such devices for some children or households.

Prior to initiating a study at Chimfunshi, it must first be submitted to the Chimfunshi Scientific Advisory Board. The board members, along with the sanctuary directress, review and determine the feasibility of the proposed study. Once the necessary formalities are sorted, an on-site assistant assists in reaching out to the parents. The study program is explained to each family expressing interest in participating in the study. Written consent is obtained for participation, which can be withdrawn by the parents at any point. Depending on the nature of the study, the recruitment process is carried out in collaboration with the school teachers.

**References and further readings:**

* https://chimfunshiwildlife.org/
* https://www.britannica.com/place/Zambia/Education
* https://legendsandlegaciesofafrica.org/sheilasiddle.php

### Hai||om (Namibia) - Ethnic group

The Hai||om are an indigenous group primarily living in the areas in and around the Etosha Pan of northern Namibia. In the current research, participants were recruited in small communities of 150 to 300 people in the Oshikoto and Otjozondjupa regions.

They live in a dry Savannah of northern Namibia, characterized by the arid Namib Desert and the fringes of the Kalahari. The climate is hot and dry, with average temperatures exceeding 30°C and unsteady and minimal rainfall during summer. Vegetation is sparse, consisting mainly of drought-resistant shrubs and grasses with bloom of edible and non-edible plants during rain season.

The Hai||om (i.e., treesleepers), also referred to as Heikum or Hai//om, speak Hai||om, which belongs to the Khoekhoe language family often recognized for containing distinct consonants (i.e., “click” sounds). Local dialects (i.e., =Ākhoe) differ across communities. They traditionally lived in small, semi-nomadic bands, loosely organized around kinship ties and shared leadership roles. Like many other forager communities of southern Africa, the Hai||om value egalitarianism, autonomy, and sharing on demand as foundational schemas organizing their social structure.

The Hai||om are the indigenous people of the Etosha pan, have been living in this area for centuries. During the colonial era and German and South-African occupation of their lands, the Hai||om faced forced relocation from Etosha to the neighboring farmlands. The dispossession of their traditional lands together with political and societal discrimination continue having lasting impact on their communities and livelihoods.

Traditionally, the Hai||om have been hunter-gatherers, relying on their vast knowledge of the desert environment to find food and water. They utilized a variety of techniques for hunting and gathering, including tracking, digging, and setting traps. They exhibited a semi-nomadic lifestyle, taking advantage of seasonal  resource availability. This self-sufficient lifestyle has been impacted by factors like climate change and government programs, but many Hai||om continue to practice traditional subsistence activities whenever possible. Today, the rural Hai||om communities participating in the current research practice a mixed subsistence combining seasonal and occasional farm work, small-scale gardening and animal husbandry, foraging, but also rely on government food programs.

Most of the Hai||om families participating in this research house in small huts constructed from branches, corrugated iron, and animal dung. These dwellings are easily portable and household mobility remains high in these communities. Household composition is flexible, with extended families often living together and individuals moving between and within  communities based on kinship ties and resource availability.

Hai||om child-rearing emphasizes autonomy, egalitarianism, cooperation, and sharing on demand as foundational schemas. Children are actively involved in daily activities from a young age, learning essential skills primarily through observation and participation, rather than direct pedagogy. Parents also instill cultural knowledge and values through storytelling and traditional practices.

Due to their historical marginalization and remote location, access to formal education remains challenging for many Hai||om communities. In one community, children have access to a primary school from age 6 onward, whereas in the other community, children need to travel about 35km to access the next school. In both communities, dropout rates are reported low by teachers, likely due to multiple reasons, such as cultural mismatches between often authoritarian teaching styles and foreign learning content at schools, children’s autonomy in deciding on their learning environments, and secondary schools being located far away from their home communities. Traditional knowledge transfer remains vital, with elders playing a significant role in childcare and education. At the same time, children spend considerable amounts of their daily lives in mixed-age peer groups, which present important learning contexts for these children. Children learn through practical experience and participation in daily activities alongside adults and peers.

Exposure to technology, particularly smartphones and other touchscreen devices, is highly limited among the Hai||om due to their remote location and traditional lifestyle. Electricity is available through solar panels at schools, farm houses, and few households, but the importance of electricity in people’s lives remains minimal. Some individuals, particularly younger generations, may have some exposure to technology through interactions with outsiders or limited access to shared devices, such as cellphones.

It is important to note that this depiction provides an overview on the communities participating in the current research and does not reflect the diversity of livelihoods among urban-dwelling Hai||om people more generally. Their experiences and adaptations vary depending on individual circumstances and ongoing socio-economic changes.

Children were recruited from their homes as part of a long term research project taking place in these communities. They were later tested in quiet locations either at the school building or vacant spaces in their communities. Testing was done by male adult experimenters and a research team known to the communities for about 8 years. The research was approved by the Max Planck Research Ethics Council, the Namibian Commission on Research Science and Technology (Permit Number RPIV01112021), the Namibian San Council, and the Ministry of Education, Arts and Culture of the Republic of Namibia. The study also adhered to guidelines for best practice in cross-cultural research (Bruno et al., 2022). Children received small amounts of candy as a compensation for their participation.

**References and further readings:**

* Bruno, D., Pope-Caldwell, S. M., Haberl, K., Hanus, D., Haun, D., Leisterer-Peoples, S., Mauritz, S., Neldner, K., Sibilsky, A., & Stengelin, R. (2022). Ethical guidelines for good practice in cross-cultural research. Leipzig: Department of Comparative Cultural Psychology, Max Planck Institute for evolutionary anthropology.

* Dieckmann, U. (2012). Born in Etosha: Living and learning in the wild. Legal Assistance Centre.

* Dieckmann, U., Thiem, M., Dirkx, E., Hays, J., Legal Assistance Centre (Namibia), & Desert Research Foundation of Namibia (Eds.). (2014). Scraping the pot: San in Namibia two decades after independence. Land, Environment and Development Project of the Legal Assistance Centre and Desert Research Foundation of Namibia.

* Friederich, R., & Lempp, H. (2014). Etosha: Hai||om heartland: ancient hunter-gatherers and their environment. Namibia Publishing House (Pty) Ltd.

* Widlok, T. (1999). Living on Mangetti: “Bushman” Autonomy and Namibian Independence. Oxford University Press.

### Khwe (Namibia) - Ethnic group

The Khwe (also referred to as Khoe, Khwedam, or Barakhoena) people reside in their ancestral lands within today’s Bwabwata National Park in northwestern Namibia and the bordering regions in Angola and Botswana. There is also a Khwe community in today’s South Africa that has moved there while Namibia became independent in 1990. The current research was conducted in the Namibian Zambezi region which is home to numerous small Khwe communities. There are an estimated 3,000 to 5,000 Khwe residing in Bwabwata as of today.

Due to its sub-tropical climate, inhabitants of Bwabwata National Park experience hot and humid summers, whereas other seasons are much more characterized by a hot and dry climate. Bwabwata is a dual-use National Park in which wild animals, including large predators, elephants, and other animals native to the Zambezi region roam freely around the Khwe and Hambukushu settlements in the Park.

The Khwe people speak Khwedam, which is part of the larger Khoe-Kwadi language spectrum. They traditionally lived in small, semi-nomadic bands with leadership based on a combination of age and experience. Today, people live in rural communities of varying size, ranging from one-household settlements to settlements of up to 600 people. Larger settlements have access to schools, medical centers, police stations, and small shops.

The Khwe of Bwabwata are the indigenous people of the region and have resided in the area for centuries. They faced similar historical injustices as other indigenous groups of Southern Africa, such as the Hai||om, resulting from colonization and political marginalization. This includes, but is not limited to, colonial dispossession, forced relocation, and the neglect of their traditional leadership systems. These continue to have lasting impacts on their community and land rights until today.

Traditionally, the Khwe lived as hunter-gatherers, relying on their extensive knowledge of the environment to hunt game, gather bushfood, and find water. This self-sufficient way of life has been impacted by various factors, including the migration of Hambukushu (Bantu) communities to the area, German colonization and South African occupation, governmental resettlement programs, and other environmental changes. Today, Khwe people are no more allowed to leave their settlements freely as part of the anti-poaching regulations of the dual-use National Park. However, traditional subsistence activities continue being an integral part of the cultural identity of many Khwe people. Their traditional subsistence is today supplemented with wage labor, seasonal harvesting programs, small-scale gardening and animal husbandry, as well as governmental and non-governmental support.

Household composition is flexible, with extended families often residing together and individuals moving between groups based on kinship ties and resource availability. Before Namibian independence, the South African Military Forces had multiple camps in the region due to its proximity to the Angolan border. After independence, Khwe communities increasingly settled in the residential infrastructure they had left behind. Other communities live in small huts made of wood, dung, corrugated iron, and other building materials.

Similar to the Hai||om, Khwe child-rearing emphasizes autonomy, egalitarianism as well as strong norms on cooperation and demand sharing. Children actively participate in daily activities from a young age, learning essential hunting, gathering, and survival skills through observation and participation. Parents also transmit cultural knowledge and values through storytelling and traditional practices.

Access to formal education has been historically limited for the Khwe of Bwabwata due to their remote location and marginalization. As of today, formalized education presents a two-edged sword as it is perceived by many as an important opportunity for social improvement, while also contributing to the erosion of cultural and environmental knowledge and skills. In the community sampled here, children do have access to a local kindergarten and a combined school. Many learners attending school did not experience kindergarten education as they grew up in the smaller communities surrounding this larger settlement. Especially in early childhood, traditional knowledge transfer and parenting practices remain crucial, with elders playing a significant role in childcare and education. Children learn through practical experience and participation in daily activities alongside adults. Additionally, recent initiatives aim to improve access to formal education within the national park.

Exposure to technology, particularly smartphones and other touchscreen devices, is generally limited among the Khwe of Bwabwata also due to their remote location, economic hardships, and lack of access to electricity. However, similar to the Hai||om, the level of access can vary depending on proximity to towns and settlements. Some individuals, especially younger generations, may have some exposure through interactions with external communities or shared devices.

It is important to note that this is only an approximate depiction of the situation faced by the communities participating in the current research. It does by no means give credit to the diversity of livelihoods among the Khwe people more generally. Their experiences and adaptations vary depending on individual circumstances and ongoing socio-economic changes.

Children were recruited for a long-term research project after consulting with community elders and representatives. The project was conducted in local schools or kindergartens and received necessary approvals from relevant ethical bodies in Namibia and Germany: The Max Planck Research Ethics Council, the Namibian Commission on Research Science and Technology (Permit Number RPIV01112021), the Namibian San Council, and the Ministry of Education, Arts and Culture of the Republic of Namibia. Self-commitment regarding research ethics aligned with departmental guidelines for best practice in cross-cultural research (Bruno et al., 2022). Testing was carried out by a male adult from the community alongside a familiar research team. The study adhered to strict ethical guidelines for cross-cultural research, and children received small tokens of appreciation for their participation.

**References and further readings:**

* Dieckmann, U., Thiem, M., Dirkx, E., Hays, J., Legal Assistance Centre (Namibia), & Desert Research Foundation of Namibia (Eds.). (2014). Scraping the pot: San in Namibia two decades after independence. Land, Environment and Development Project of the Legal Assistance Centre and Desert Research Foundation of Namibia.

* Nkengbeza, D., Kangumu, B., & Sibuku, C. (2021). The Perceptions of the KHWE (San) Community at Bwabwata on the Importance and Influence of Formal Education (SSRN Scholarly Paper 3886443). https://papers.ssrn.com/abstract=3886443

* Koot, S., Beek, W. V., & Diemer, J. (2016). The Khwe of Namibia. Foragers between Game, Tourism, and Politics. Anthropos, 111(2), 497–512. https://doi.org/10.5771/0257-9774-2016-2-497

* Paksi, A. (2019). ‘We live in a modern time’: : Local perceptions of traditional knowledge and formal education among a Namibian San community. Hunter Gatherer Research, 5(1–2), 119–151. https://doi.org/10.3828/hgr.2019.6

* Paksi, A. (2020). ‘Surviving Development’: Rural development interventions, protected area management, and formal education with the Khwe San in Bwabwata National Park, Namibia. Faculty of Social Sciences, University of Helsinki.

### Leipzig (Germany) - Location

Leipzig is a medium-sized city located in the east of Germany with a population of approximately 620,000. It serves as a significant urban center within the state of Saxony. Geographically, Leipzig is situated in the southernmost part of the North German Plain.

The city landscape is characterized by urban and suburban areas, with parks, squares, and historical buildings. Leipzig has an oceanic climate with continental influence. Winters are cold (average temperature around 1°C) and summers warm (average temperature around 19°C).

The majority of inhabitants of Leipzig are German, many of which have moved from other parts of Germany. Around 17% of the population have a migration background. The largest groups of migrants come from Syria, Ukraine, and the Russian Federation. In comparison to more rural areas in the region, Leipzig has a larger proportion of inhabitants with a migration background and is generally seen as more liberal and progressive. German is the primary language spoken in Leipzig. Additionally, due to its role as a major educational and cultural center, English and other foreign languages are also spoken by a significant portion of the population.

Historically, Leipzig gained prominence as a trade city during the Middle Ages and later became a major center of the Enlightenment and the arts in the 18th and 19th centuries. Following World War II, Leipzig was part of the socialist German Democratic Republic. Leipzig played a crucial role in the Peaceful Revolution of 1989, which ultimately led to the fall of the Berlin Wall and the reunification of Germany.

Leipzig has been a market economy for centuries. It has a modern and diversified economy that encompasses various sectors, such as manufacturing, services, and creative industries. The city is known for its trade fairs, music, and publishing industries.

The city has a combination of apartment buildings, detached houses, and residential complexes. Household composition varies, ranging from single individuals and couples to families with children and multi-generational households. Children mostly live in nuclear families with one or two children and their caregivers.

Parental socialization goals in Leipzig align with broader German cultural norms and typically emphasize the importance of children’s independence, psychological autonomy, self-confidence and assertiveness. Education and learning are seen as important aspects of children’s lives. Parental education levels are generally very high; adults have at least nine years of formal schooling and high school and university degrees are very widespread.

Parents spend large parts of the day away from home, and the majority of children are cared for externally: starting at nine months of age, children can attend nurseries, moving on to kindergarten at three and primary school around six. Institutional childcare is provided by the city and financially subsidized; institutionalized primary education is free and compulsory.

Children growing up in Leipzig have a high exposure to technology. Smartphones and other touchscreen devices are prevalent in most families and educational institutions, reflecting the widespread use of digital technology for communication, entertainment, and information retrieval.

Children were recruited via a database of children whose parents expressed interest in participating in studies of child development. They were tested in a separate room in their kindergartens by a German-speaking research assistant.

**References and further readings:**

* City website for demographic information: https://english.Leipzig.de
* Otto, H., & Keller, H. (2016). Is there something like German parenting. Contemporary Parenting: a Global Perspective, 81-94.

### Malatya (Turkey) - Location

Malatya is a big city located in the Eastern Anatolia region of Turkey with a population of approximately 800,000. Due to its strong economy based on agriculture, it is one of the biggest cities in eastern Turkey.

The city landscape is characterized by mostly urban and some rural areas. Malatya has a temperate continental climate with very cold winters (average temperature around -3°C) and very hot summers (average temperature around 33°C).

The population of Malatya is diverse which include inhabitants with Turkish, Kurdish, Circassian, Armenian, and Zaza origins. Malatya is also a multilingual city. Although the dominant language is Turkish, around 40% of the population speaks Kurdish.

Malatya has a long history since the Hittite times. It has also been significant location for the Roman Empire, Byzantine Empire, and Ottoman Empire.

Malatya’s economy depends on agriculture, and it is famous for apricot production. Recently, Malatya has also undergone some industrial development such as textile manufacturing.

The urban area of Malatya consists of apartment buildings. Household composition usually involves families with children and multi-generational households. Children live both in nuclear families (with parents and siblings) and in extended families (nuclear family and a set of grandparents). Parental socialization in Malatya aligns with broader Turkish cultural norms, and typically emphasize the importance of children’s interdependence, value of family. Education and learning are seen as important aspects of children’s lives. Parental education levels are generally above the average of Turkey; with most adults have high school or university degrees.

Fathers spend large parts of the day away from home, and the majority of children are cared at home by their mothers and grandparents until they start formal schooling at the age 7. Primary education is free and compulsory.

Children growing up in Malatya have some exposure to technology. Smartphones and other touchscreen devices are somewhat prevalent in most families, reflecting the widespread use of digital technology for communication.

Children were recruited through nurseries and schools who expressed interest in participating research. They were tested in a separate room in their nurseries/school kindergartens by a Turkish-speaking research assistant. Some children were also tested in parks and outdoor areas.

### Nyabyeya (Uganda) - Location

Nyabyeya Forestry College is situated in the Masindi district of western Uganda, approximately 25 km from Masindi Town and 240 km from Kampala, the capital city. The college is adjacent to Budongo Tropical High Forest and 8 km from Budongo conservation field station, where primatologists study wild-ranging chimpanzees. The college draws its sample from the surrounding villages, including Marram, Kyempunu, Nyabigoma, Nyabyeya 1 and 2 and Trading Center, Nyakafunjo, and Kyamasuka.

Nyabyeya is located on a green, gently sloping area behind Nyabyeya Hill and is characterized by tropical vegetation and a savanna climate, characterized by daytime temperatures reaching 32°C/90°F and cooler nights at 17°C/63°F. The region observes two distinct agricultural seasons - a wet season from late March to June and August to November, followed by a dry season from early December to March.

The studied villages are culturally very diverse and house ethnic groups such as Bantu, Luo, Nilotics, and Nilo-Hamites. Spoken languages depend on the tribe and include Alur, Lugbara, Runyoro, lendu, kakwa, Acholi, Langi, Luganda, Logo, Kebu, madi, and many others. Kiswahili and English often serve as a common ground. For this study, children who understand Kiswahili were invited.

During the Second World War, many Polish refugees were brought to Uganda and located in Bunyoro on the fringes of Budongo Forest at Nyabyeya. When the refugees were evacuated after World War II ended, the forestry college was transferred into the facilities the Polish refugees had left behind. It now serves as an educational institute conducting training in Forestry, Agroforestry, Biomass Energy Technologies, and beekeeping.

The community predominantly practices subsistence farming, cultivating crops like maize, cassava, sweet potatoes, and beans. Wealthier families sometimes keep goats, pigs, chickens, or ducks. Additionally, sugarcane and coffee cultivation contribute significantly to the local economy, with the Kinyara Sugar Company playing a pivotal role in sugarcane harvesting.

Living conditions in the community are characterized by extended families residing together, with an average of 10 people under one roof. While temporary round mud huts are common, permanent houses are found in the trading center of Nyabyeya. Most people in the community live by hand to mouth, and parents (especially women) work hard to provide the basics for their household members.

Parental socialization goals vary, with limited guidance for children in some instances, leading to very independent lives at a young age. Cultural practices include the belief in producing many children as a sign of respect, with specific tribes attaching wealth to the number of daughters. Twins are highly honored, and rituals are performed to welcome them. Often, young children take care of their siblings and carry them around on their backs. Children start working early, for example, by helping their parents with farming or cow herding. Often, families cannot afford school materials for their children and struggle to provide regular food. Early marriages and teenage pregnancies often result in girls leaving school earlier.

Parental education levels are generally low, with many parents above 50 having only primary education. The education system faces challenges such as inadequate school supplies, leading to high dropout rates. Children start primary school at ages 5-7, and secondary education runs for four years, with options to pursue advanced levels.

The community exhibits low exposure to technology, with smartphones and touchscreen devices considered luxuries. Limited internet usage and power supply challenges contribute to this low adoption. However, technology is more prevalent in institutions like Nyabyeya Forestry College.
Participating families were recruited by the local research assistants, for example, via personal contact, visiting them at home, or calling them. Families were invited to the Nyabyeya Forestry College (transport offered via boda boda (motorcycle)), and children participated in a quiet room in the College.

**References and further readings:**

* Budongo Conservation Field Station: http://www.budongo.org/
* Holden, E., Buryn-Weitzel, J. C., Atim, S., Biroch, H., Donnellan, E., Graham, K. E., Hoffman, M., Jurua, M., Knapper, C. V., Lahiff, N. J., Marshall, S., Paricia, J., Tusiime, F., Wilke, C., Majid, A., & Slocombe, K. E. (2022). Maternal attitudes and behaviours differentially shape infant early life experience: A cross cultural study. PLOS ONE, 17(12), e0278378. https://doi.org/10.1371/journal.pone.0278378

### Ocuilan (Mexico) - Location

Ocuilan is a municipality located in the centre south of Mexico with a population of 36,223 inhabitants, distributed in 48 towns in an area of 344.84 km2  inside the Mexican State (Ocuilan, n.d.). Geographically Ocuilan has an altitude of between 1000 to 2,600 m situated 80 km south of Mexico City. (INEGI,2010)

The communities in Ocuilan are distributed on a mountain range called Trans-Mexican volcanic Belt surrounded by valleys and mountains, the climate is semi-humid with summer rains, and average temperatures vary between 8C and 22C throughout the year. The higher elevations have mixed pine forests with more tropical vegetation in the lowlands. The distribution of the land is 31.91% agriculture, 0.48% urban zone and the rest forest. (INEGI,2010)

Historically, the municipality of Ocuilan was occupied by three different towns: Tlahuicas, Mazahuas, and Ocuiltecos. Besides being in the same territory, they had their own language, religion, and customs until the arrival of the Mexica people. Within 100 years, the Mexica dominated the area. The territory was under Mexican control until the arrival of the Spanish kingdom, which took advantage of the discontent of the towns dominated under the Aztec Empire. The Spanish kingdom allied itself with those towns and waged war on the Aztec Empire, succeeding in taking the Mexica capital on 13 August 1521. Although there was some resistance, all the towns within what is now the State of Mexico fell to the Spaniards shortly thereafter. (Edo Mex, 2018)

During the period of Spanish control, the municipality of  Ocuilan was primarily characterized by agricultural activities. The presence of missionary friars played an important role in the dissemination of catholic religion and the Spanish language. (Edo Mex, 2018)  Over the course of the conquest, the Spanish language gradually gained prominence, but it is still possible to find indigenous languages like Tlahuica, Nahualt and Mazahua among the inhabitants. (Ocuilan, n.d.)

Following the attainment of Mexican independence, Ocuilan continued to be a farming area and also emerged as an important religious centre, attracting pilgrims from various regions of Mexico. During the Mexican Revolution, the locality assumed significance as a stronghold of the Zapatistas which played an essential role in the enactment of the Political Constitution of 1917.  In the subsequent decades, political stability, economic development, and population growth in the State of Mexico. (Edo Mex, 2018)   Nowadays, livestock farming and trout rearing stand as the dominant economic pursuits in Ocuilan. (Ocuilan, n.d.)

Demographically, the average household in Ocuilan comprises 4.1 inhabitants, (Ocuilan, n.d.)  primarily adhering to a nuclear family structure characterized by an authoritative rearing model. Parents assume a proactive role in establishing parental control, emphasizing the enforcement of rules and disciplinary measures (Delfín-Ruiz et al., 2021). The parent's education level generally extends to the middle school level, while the prevailing family values involve the mother's engagement in domestic responsibilities and the father's participation in occupational pursuits.

Regarding education, children in Ocuilan typically start their day by attending kindergarten between 8 am to 12 pm. Afterwards, mothers or other family members take care of the children as formal childcare options are limited. Consequently, children typically begin their early schooling experience at around 3 years of age.

In terms of technology, approximately  77.9% of households possess at least one screen device, whereas access to the Internet is available to only  22.7% of households. Consequently, while a significant proportion of children have experienced digital technology a considerable segment lacks exposure and accessibility to such resources. (Ocuilan, n.d.)

For the study, we first obtained permission from Mexican school authorities to visit local kindergartens. Next, we contacted parents and invited them to take part in the research. The evaluation of interested children was then carried out in a separate area within their own kindergartens, overseen by a Spanish-speaking research assistant.

**References and further readings:**

* Compendio de información geográfica municipal 2010. Instituto Nacional de Estadística, Geografía e Informática INEGI (2010).  Retrieved 3 August  2023, from https://en.www.inegi.org.mx/app/biblioteca/ficha.html?upc=702825293031
* Delfín-Ruiz, C., Orozco, C. S., & Guzmán, R. C. (2021). Caracterización de los roles familiares y su impacto en las familias de México. Revista de Ciencias Sociales, 27.
* Ocuilan: Economy, employment, equity, quality of life, education, health and public safety. (n.d.). Data México. Retrieved 13 May 2023, from https://datamexico.org/en/profile/geo/ocuilan
* Historia del Estado de Mexico. (2018) Gobierno del Estado de México Edo Mex (2018).  Retrieved 13 May 2023  from  https://edomex.gob.mx/historia_edomex

### Plymouth (UK) - Location

Plymouth is a coastal city in the South West of England with a population of more than 260,000 inhabitants. About 91% of the population was born in the UK and of those born outside of the UK, the largest proportion comes from the EU (4.5%). 95% of residents identified as White, 2.3% as Asian, and 1% as Black. English is the predominant language spoken in the city.

The city has a temperate oceanic climate with average temperatures ranging from 4 to 20 degrees Celsius. Winters are generally mild, and temperatures rarely drop below zero degrees. Rain is frequent and highest in autumn and winter.

The area was first settled in the Bronze Age. Plymouth is situated at the meeting point of two rivers (river Tamar, river Plym) and has a large natural harbour. Due to its location, the city has served as a shipping port and naval harbour throughout history. It played an important role in the Spanish-British war in the 16th century and was a departure port for settlers bound for North America in the early 17th century. Due to its importance as a naval port, the city was heavily bombed during the 2nd World War and most of the city center needed to be re-build post-war.

In the present day, Plymouth is home to a major dockyard and the biggest naval base in Western Europe. The largest employers are the defence industry and the armed forces, the university, the national health service as well as maritime businesses. Tourism, particularly in the summer months, is another source of income. The city has some areas of very high deprivation (among the 1% of most deprived areas in England) and the median income (33k) is slightly below the UK median (35k).

The city has a large waterfront area and several harbours and docks, and there is a large inner-city park. Housing in the city mostly consists of terraced and semi-detached houses with few multi-storey buildings.  
Children mostly grow up in nuclear families. Children in the current study came mostly from middle class families. In Western middle-class families care-giving is usually child-centered and parents are responsive to children’s needs and wishes. Autonomy and independence are valued but, compared to countries such as Germany, children are somewhat less independent. As the UK is a multicultural country (and multiculturalism is increasing), parenting values are heterogenous—though the population in Plymouth, as compared to the rest of the UK, is less heterogenous and predominantly White British.

The UK has one of the highest childcare costs in Europe (52% of median female earnings). Nurseries will offer places for children from 3 months of age, but currently the government only subsidizes childcare from three years of age (caregivers receive 15 hours of free nursery care for three- and four-year-olds). Primary school starts at age 5 years, and secondary school at age 11 to 12 and lasts until 18 years. Education is compulsory in the UK, but children do not have to attend school and can be homeschooled instead.

Children are exposed to technology from early on, including mobile phones and other touch screen devices.
Children were mostly recruited through a local primary school and some additional children were recruited through personal contacts of one of the experimenters. Parents received information letters and consent forms. Experimenters also visited the school during drop off times in the morning to answer any questions parents may have had. Only children whose parents had consented to their participation took part in the study. In the school setting, two experimenters usually conducted the study in parallel in quiet rooms in the school. Some children who were recruited through personal contacts participated in the study online, facilitated by one experimenter.

**References and further readings:**

* Bozicevic, L., Hill, J., Chandra, P. S., Omirou, A., Holla, C., Wright, N., & Sharp, H. (2023). Cross-cultural differences in early caregiving: levels of mind-mindedness and instruction in UK and India. Frontiers in Child and Adolescent Psychiatry, 2, 1124883-1124883.

* Kagitcibasi, C. (2017). Family, self, and human development across cultures: Theory and applications. Routledge.

* Pearson, E., & Rao, N. (2003). Socialization goals, parenting practices, and peer competence in Chinese and English preschoolers. Early Child Development and Care, 173(1), 131-146.

* Plymouth City Council, Understanding Our Communities. Retrieved from: https://www.plymouth.gov.uk/understanding-our-communities

* Mubashar, L. (2023). 'Childcare puzzle': Which countries in Europe have the highest and lowest childcare costs? Retrieved from: https://www.euronews.com/next/2023/03/06/childcare-puzzle-which-countries-in-europe-have-the-highest-and-lowest-childcare-costs

### Pune (India) - Location

Pune is located in the state of Maharashtra in western India. The city’s total area is 15.642 square kilometres and the municipal corporation area covers 518 square kilometers. It lies on the western margin of the Deccan plateau at an altitude of 1840 feet. It is on the leeward side of the Sahyadri mountain range which forms a barrier with the Arabian Sea. The population within city limits is 4.5 million while including the metro population, it stands at 7.1 million. Pune is one of the largest information technology hubs in India and is also an important automobile and manufacturing hub. Sixty-two percent of the population in Pune is under the age of 30. The top three industries that offer employment are automotive, manufacturing and information technology and start-ups. Automobile industries such as Tata Motors, Jaguar, Renault, Volkswagen, General motors, Skoda, Force motors, Kinetic motors  and Mercedes-Benz have their manufacturing plants in Chakan, an industrial area around Pune. Cummins, a leading manufacturer of diesel and natural gas engines as well as JCB, the farm equipment company are other renowned investors. The manufacturing industry provides employment to over 600,000 people while the IT sector employs over 350,000 people. About 2.2 million migrant workers are employed in the automobile, construction and medium scale manufacturing industries.

Pune is referred to as the ‘Oxford of the East’ because of its reputed educational institutions and is also considered to be one of the most livable cities in India. The literacy rate in Pune is 86.15%.  As a result of rapid industrialization since the 1960s, a large number of people moved to Pune. Since there was inadequate housing to meet the demand, there was an increase in the number of slum dwellers. Nearly half of the slum dwellers do not have in-house toilets and 10% do not have electricity. 
Pune experiences a tropical wet and dry climate with average temperatures ranging between 20 and 28 degrees Celsius. Summer, monsoon and winter are the three main seasons with May being the hottest month. During summer, i.e. mid-March to mid-June, temperatures can go as high as 42 degrees Celsius. Monsoon lasts from June to October with most of the annual rainfall of 722mm falling between June and September with July being the wettest month. In the months of December and January, temperatures remain around 29 degrees Celsius with overnight temperatures dropping to below 12 degrees.

During the 18th century, Pune was part of the Maratha empire and seat of the Peshwas. Besides it has been ruled by dynasties such as the Rashtrakutas, Ahmednagar Sultanate, Mughals and Adil Shahi dynasty.  The modern city of Pune is comprised of many distinct neighbourhoods. These include ‘peths’ [Marathi for locality] on the eastern bank of the Mutha river and the cantonment areas of Khadki and Camp. Pune is considered to be the cultural capital of Maharashtra with emphasis on education, arts and crafts, music and theater. The most common language spoken is Marathi. Hinduism is the major religion practiced by just under 80% of the population. Other religions practised are Islam, Buddhism, Jainism, Christianity, Sikhism and Zoroastrianism. Food is made of sorghum and pearl millets with a fairly large consumption of coconut, garlic and green chillies to spice up the food.

Children in Pune normally enter school between the ages of 2.5 and 3.5. The school leaving exam is taken between the ages of 17 and 18. Grades 11th and 12th (equivalent to AS and A levels in the UK) may be attended in school or ‘junior college’ with the latter offering education up to the postgraduate level. Following a school day, it is not uncommon for children to attend tuition or focussed coaching in sports or arts. Most households will have at least one mobile phone as well as a laptop so children are exposed to technology fairly early.

Children are taught to address their elders (either in English or the equivalent term in a local language) as ‘older brother/sister’, ‘uncle’, ‘aunty’ or ‘grandmother’/’grandfather’ even if they are not relatives. This is considered to be respectful. At times, children touch the feet of elders as a sign of respect. Parents encourage their children to visit family and friends particularly during festivals. 
For the current study, recruitment took place via personal contacts. Both schools have participated in previous research projects and the researcher is on the management committee of one of the schools. The researcher was given a separate quiet room to carry out testing. Normally, after one child finished the task, he or she would send their classmate who had volunteered for the project. In the case of very young children, the researcher would the children personally from their respective class and ensure that they were handed back to their teachers upon completion.

**References and further readings:**

* https://en.wikipedia.org/wiki/Pune
* https://www.iimb.ac.in/turn_turn/pune-industrial-workforce.php
* https://pune.gov.in/culture-heritage/
* https://www.bbc.com/travel/article/20181028-an-indian-city-with-audacious-attitude

### Stanford (USA) - Location

The sample was drawn from a children’s museum in San Jose, California located in the western United States of America. San Jose is a large city with a population of approximately 983,000 people and geographically is within the Santa Clara Valley in the southern part of the Bay Area, California.

The city is a large metropolitan area with urban and suburban areas bordered by mountain ranges to the east and west. The land is characterized by bay flats and hills with several rivers running through the city. San Jose’s climate is described as mild, semi-arid, and Mediterranean. A wet season typically occurs between November and March with combined rainfall and an average daytime temperature of 50 degrees Fahrenheit. A dry season occurs between March and November where precipitation is minimal and the average daytime temperature is 70 degrees Fahrenheit.

According to the most recent census data, the 3 largest groups by Race are Asian (Non-Hispanic) (37%), White (Non-hispanic) (25%) and Hispanic (31%). 40% of residents immigrated to the United States, with most residents originating from Mexico, Philippines, and China. Although English is the primary language used, more than half (58%) of people also speak a language other than English at home, including Spanish (20%), Chinese (7%), Tagalog (3%), Vietnamese (2%), and other Asian & Pacific Islander languages (19%).

San Jose was originally home to the Tamien people of the Ohlone from around 4000 BC. Around 1777 the Tamien people were displaced and enslaved, and San Jose became the first town in the Spanish colony Nueva California. San Jose would eventually become the first incorporated city and state capital when California became a state in the United States of America following the Mexican-American War.

San Jose historically is a large farming community with a major industry in food processing and agriculture. The Valley was especially known for its tree nuts, berries, and orchard fruits. The city soon became a hub for the manufacturing of electric machinery and motor vehicles, and would eventually become the modern technological hub that it is today. San Jose is presently centered in “Silicon Valley”, a global center for technological innovation and entrepreneurs.

The city has a combination of apartment buildings, detached houses, and residential complexes. Household composition varies, ranging from single individuals and couples to families with children and multi-generational households. Children mostly live in nuclear families with one or two children and their caregivers.

Parental socialization goals in California typically emphasize tolerance, individualism, diversity, personal fulfillment, and holistic autonomy, but vary greatly as a result of its diverse populous. Education and learning are seen as important aspects of children’s lives. Parental education levels are generally very high; many adults have a graduate degree, and most adults have at least 12 years of formal schooling and a 4-year college degree.

It is common for parents to spend large parts of the day away from home, and typically rely on external care including family, nannies, and daycare for childcare. Daycares are largely commercialized and non-subsidized. Starting around age 2, children can begin attending preschool. At age 5, children can enter kindergarten in publicly subsidized schools and move on to primary school at age 6. While public primary education is typically free and compulsory, around 15% of children in San Jose are enrolled in private, non-subsidized institutions for primary school.

Children in San Jose have high exposure to technology. Nearly all households in San Jose have a computer (97%), with 93% of households having a broadband internet subscription. Smartphones and other touchscreen devices are used in most families and are used widely for education, communication, and entertainment.

Children were recruited in person at a children’s museum in San Jose. Participants were randomly solicited from patrons visiting the children’s museum. Parents were given information about the study activity and consented before beginning the study. Children completed the study in a separate research room centrally located in the museum.

**References and further readings:**

* Miller, Miguel (1999). Climate of San Jose, NOAA Technical Memorandum, retrieved 22 September 2023 from: https://www.wrh.noaa.gov/mtr/sfd_sjc_climate/sjc/SJC_CLIMATE3.php#:~:text=San%20Jose%20%3Ds%20latitude%20and,total%20falls%20within%20this%20period.
* Quick Facts: San Jose City, California (2022). United States Census Bureau, retrieved 22 September 2023 from: https://www.census.gov/quickfacts/fact/table/sanjosecitycalifornia/PST045222
* San Jose, California Population 2023 (2023). World Population Review, retrieved 22 September 2023 from: https://worldpopulationreview.com/us-cities/san-jose-ca-population
* San Jose Demographics and Diversity (2023). San Jose, retrieved 22 September 2023 from: https://www.sanjose.org/meetings/quick-guides/san-jose-demographics-and-diversity
* Wilcox, W. D., Wang, W. (2020). State of Contradiction: Progressive Family Culture, Traditional Family Structure in California, Institute for Family Studies, retrieved 22 September 2023 from:
https://ifstudies.org/blog/state-of-contradiction-progressive-family-culture-traditional-family-structure-in-california

### Windhoek (Namibia) - Location

**References and further readings:**

* ...

```{r figpred2, include = T, fig.align = "center", fig.cap = "Household demographics by community.  Points show individual data points with minimal vertical and horizontal jitter to avoid overplotting. Red points show means with 95\\% CI.", out.width="100%"}

# dempred_agg <- data%>%
#   group_by(community)%>%
#   distinct(subjid, .keep_all = T)%>%
#   select(subjid, community, continent, children, younger_children, household)%>%
#   pivot_longer(names_to = "dem", values_to = "value", cols = c(children, younger_children, household))%>%
#     mutate(dem = recode(dem,
#                       younger_children = "younger children"),
#          dem = factor(dem, levels = c("household", "children", "younger children")))%>%
#   group_by(community, dem)%>%
#   na.omit()%>%
#   tidyboot_mean(col = value)
#   
# 
# dempred <- data%>%
#   group_by(community)%>%
#   distinct(subjid, .keep_all = T)%>%
#   select(subjid, community, continent, children, younger_children, household)%>%
#   pivot_longer(names_to = "dem", values_to = "value", cols = c(children, younger_children, household))%>%
#   mutate(dem = recode(dem,
#                       younger_children = "younger children"),
#          dem = factor(dem, levels = c("household", "children", "younger children")))%>%
#   ggplot(aes(y = value, x = community))+
#   facet_grid(dem~., scales = "free", space = "free_x")+
#   geom_jitter(width = 0.2, height = 0.1, alpha = 0.1)+
#   geom_pointrange(data = dempred_agg, aes(ymin = ci_lower, ymax = ci_upper, y = mean), col = "firebrick")+
#   theme_few()+
#   labs(y = "No. of individuals", x= "")+
#   theme(axis.text.x = element_text(angle = 55, vjust = 1, hjust=1))
# 
# ggsave(plot = dempred, "../visuals/dempred.png", height = 6, width = 10, scale = 1)

knitr::include_graphics("../visuals/dempred.png")

```

# Method

We adapted the task developed by Prein and colleagues [@prein2023tango]. We refer to the original publication for a detailed description of its development, implementation and psychometric evaluation (in Germany). Below give an overview and focus on the cross-cultural adaptation of the task. The task itself, including all the versions used in the study, can be accessed via the following website: https://ccp-odc.eva.mpg.de/tango-cc/

## Setup and Procedure

```{r task, include = T, fig.align = "center", fig.cap = "Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was.", out.width="80%"}
knitr::include_graphics("../visuals/task.png")
```

The task was implemented as a browser-based interactive picture book using `HTML` and `JavaScript`. Participants saw animated agents on a touch screen device, listened to pre-recorded audio instructions and responded by touching the screen. In all communities, a research assistant, fluent in the local language(s), guided the child through the task.

Figure \@ref(fig:task) shows a screenshot from the task. The task was introduced verbally by the assistant as the balloon game in which the participant would play with other children to find a balloon. On each trial, participants saw an agent located in a window in the center of the screen. A balloon fell down from its starting position just below the agent. The agent's gaze followed the trajectory of the balloon. That is, the pupils and the iris were programmed to align with the center of the balloon. Once the balloon had landed on the ground, the agent was instructed to locate it, that is, to touch the location on the screen where they thought the balloon was. On each trial, we recorded the exact x-coordinate of the participant's touch.

There were two types of familiarization trials. In fam1 trials, the balloon fell down and landed in plain sight. Participants simply had to touch the visible balloon. In fam2 trials, the trajectory of the balloon was visible but it landed behind a small barrier (usually a hedge - see Figure \@ref(fig:task)). Thus, participants needed to touch the hedge where they saw the balloon land. Next came test trials. Here, the barrier moved up and covered the balloons trajectory. That is, participants only saw the agents eyes move, but not the balloon. They had to infer the location of the balloon based on the agent's gaze direction. During fam1, fam2 and the first test trial, children heard voice overs commenting what happened on the screen. Critically, the agent was described as wanting to help the child and to always look at the balloon (see section [script](#appendix-script) for the English wording).

Children completed one fam1 trial, two fam2 trials and 16 test trials. We excluded the first test trial from the analysis because of the voice-over. Thus, 15 test trials were used in the analysis below.

Each child saw eight different agents, four male, four female. The agent changed from trial to trail, with alternating genders. A coin toss before the first trial decided whether the first agent was male or female. The order in which agents were shown was randomized with the constraint that all agents had to be shown once until an agent was shown again. The color of the balloon also changed from trial to trial in a random order, also with the constraint that all colors appeared once before any one was repeated.

The location (x-coordinate) where the balloon landed was determined in the following way: The screen was divided in ten equally sized bins. On each trial, one of the bins was randomly selected and the exact x-coordinate was randomly chosen within that bin. Constraints were that the balloon landed in each bin equally often and the same bin appeared no more than twice in a row.

All children were tested with a touch screen device with a size between 11 and 13 inch equipped with a webcam. The data was either stored locally or sent to a server. In addition to the behavioral data, we stored the webcam recording of the session for verification purposes.

## Culture-specific adaptations

The following components of the task were adapted to each cultural settings: voice-overs, agents, background. The voice overs were originally formulated in British English and German (see section [script](#script)). For each setting, they were first translated/adapted to the local language. In some cases (BaYaka and Bandongo) this was done by first translating them to French. The translations were done by native or near-native speakers. They were not exact translations of the original text but were adapted so that children from the respective community would understand them. For example, the word "hedge" in the British English version  was changed to "bush" for the Nigerian English version to adjust for how children in the respective community would refer to the thing they saw. The content of each utterance was comparable across languages.

The agents were drawn for each cultural setting in close collaboration with local researchers and/or research assistants. They were meant to show young adults and represent the diversity in appearance prevalent in the local population. Figure \@ref(fig:figagents) shows the agents that were used in the study. Some agents were used in multiple cultural settings. All agents had the same underlying eye-geometry (size of the pupil and the iris) and only differed in superficial features.

```{r figagents, include = T, fig.align = "center", fig.cap = "Agents used in the task. Some agents were used for multiple cultural settings", out.width="100%"}
knitr::include_graphics("../visuals/agents.png")
```

## Questionnaire

The parents of children who participated in the task were asked to fill out a short questionnaire with the following questions (English version):

* Is there a device with a screen in your household? (answer: "yes" or "no")
* Does your child have experience with touchscreens? (answer: "yes" or "no")
* How many people live in your household (people residing and sleeping in one place e.g., a house)? (answer: number)
* How many children live in your household? (answer: number)
* How many children in your household are younger than the child participating in the study? (answer: number)

The questions were translated to parents native language by a native speaker. In The number of older children living in the same household was computed by subtracting one (the child taking the test) and the number of younger children from the total number of children. Whenever parents were not able to read or write (most relevant for Chimfunshi, BaYaka, Bandongo, Khwe, Ha||om), research assistants asked the question and recorded the answer.

# Analysis

We analysed the data in three steps. First, we tested whether there was substantial cross-cultural variation in performance. Second, we looked for signatures in the data that children processed gaze in the same way across cultural settings. Finally, we related performance in the task (on an individual level) to aspects of everyday experience, in particular the exposure to screen-based technology and social interaction (approximated by household composition).

We used Bayesian Regression models fit in `R` [@r] using the package `brms` [@burkner2017brms] for all analyses except the cognitive models, which were implemented in the probabilistic programming language `webppl` [@dippl]. For regression models, we used the default priors built in to `brms`. The dependent variable in all regression models was imprecision, that is, the absolute distance between the true location of the balloon (to x-coordinate of its center) and the location where the participant touched the screen. We used a Log-normal distribution to model the data because the natural lower bound for imprecision is zero and the data was right skewed with a long tail. Numeric predictors that entered the models were scaled to have a mean of zero and a standard deviation of 1. For each model we collected 6000 samples from four chains, removing the first 2000 for burn-in. All models had converging chains (Rhat values < 1.01). Details specific to each analysis can be found below. The analysis code can be found in the online repository associated with the study (link).

We pre-registered an analysis plan at https://osf.io/tdsvc. The analyses reported below follow the same logic as the pre-registered analyses but they differ substantially in detail. The main difference lies in the random effects structure we used in the models. In the pre-registration, we did not include random slopes for target centrality (i.e. the distance of the location where the balloon landed from the center of the screen), neither within subject nor within cultural setting. Instead, we included random slopes for trial within subject. We decided to change this and remove random slopes for trial but include them for target centrality. The reasons were as follows: trail effects in the sense of learning across trials were unlikely because participants did not receive differential feedback. Furthermore, in multiple studies with the same task since the registration we never found any trial effects [@prein2023tango; @prein2023variation]. We therefore removed trial from the models completely. On the other hand, we included random slopes for target centrality to be able to look at cross-cultural variation. The cognitive model was not mentioned in the pre-registration because it is a more recent development [see @prein2023variation].

0## Cross-cultural variation

As a first step, we investigated cross-cultural variation in performance. As stated in the pre-registration, this can be problematic when including random intercepts for participant and cultural setting at the same time. Participants are fully nested within cultural setting. If there was an effect of cultural setting, we would expect participant random intercepts to cluster by cultural setting. This clustering would appear whether or not cultural setting would be included in the model as a random effect or not -- the only difference would be if the participant random intercepts are estimated as a deviation from a grand intercept or a culture-specific one. Standard metrics such as `WAIC` or `LOO` would penalize the model with additional intercept for cultural setting for having additional parameters that do not help to improve predictive accuracy.

To get around this problem, we used a cross-validation procedure [see e.g., @stengelin2023children]. For each cultural setting we randomly sampled a data set that was 5/6 the size of the full data set (training data). Then we fit the model to this training data and used the estimated model parameters to predict the remaining 1/6 of the data (testing data). We then compared the model predictions from the different models by computing the mean difference between the true and predicted imprecision, over all trials in the testing data set. This approach gets around the problem mentioned in above because the model predicts a new dataset for which the individual random intercepts are unknown. Clustering by culture could therefore only be predicted by a model that included culture as a predictor. We repeated the cross-validation procedure 100 times and counted which model performed best most often.

We compared the following models (`brms` notation):

`null`:

* `imprecision ~ age + target_centrality + (target_centrality | participant)`

`culture`:

* `imprecision ~ age + target_centrality + (target_centrality | participant) + (target_centrality | culture)`

`culture_age`:

* `imprecision ~ age + target_centrality + (target_centrality | participant) + (age + target_centrality | culture)`

```{r}
# mdata <- data%>%
#   ungroup()%>%
#   mutate(ageinyears = as.numeric(scale(ageinyears)),
#          targetcentralityx = as.numeric(scale(targetcentralityx)),
#          click = abs(clickdistfromtargetcenterx))

#write_csv(mdata,"../data/mdata.csv")
```

```{r}
# bm_null <- brm(click ~ ageinyears + targetcentralityx + (targetcentralityx|subjid),
#                               data   = mdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 6000,
#                               chains = 4,
#                               cores  = 4,
#                               threads = threading(8),
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr")
# 
# saveRDS(bm_null, "../saves/bm_cv_null.rds")
# 
# 
# bm_cult <- brm(click ~ ageinyears + targetcentralityx + (targetcentralityx|subjid)+ (targetcentralityx |community),
#                               data   = mdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 6000,
#                               chains = 4,
#                               cores  = 4,
#                               threads = threading(8),
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr")
# 
# saveRDS(bm_cult, "../saves/bm_cv_cult.rds")
# 
# bm_cult_age <- brm(click ~ ageinyears + targetcentralityx + (targetcentralityx|subjid)+ (ageinyears + targetcentralityx |community),
#                               data   = mdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 6000,
#                               chains = 4,
#                               cores  = 4,
#                               threads = threading(8),
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr")
# 
# saveRDS(bm_cult_age, "../saves/bm_cv_cult_age.rds")
# 
# fixef(bm_cult_age)%>%as_tibble(rownames = "term")%>%mutate_if(is.numeric, round, 2)%>%saveRDS(.,"../saves/cult_age_fix.rds")

```

```{r}
# script for cross validation can be found in  `../scripts/cross_validation.R`
# cross validation is run in script `../scripts/run_cross_validation.R`, here we only load the saved output
cv_cult <- readRDS("../saves/cross_validation_result.rds")
```

We pre-registered the inclusion of additional models which did not include age as a predictor. To reduce computational demands and given that we did not implement this approach to determine the effect of age, we decided to leave these models out so that all models included age as a predictor. The main questions were, therefore, whether there were systematic absolute (model `culture`) or also developmental (model `culture_age`) differences between cultural settings.

The results showed that the `culture_age` model made, by far, the best predictions and outperformed the `culture` model in `r round(mean(cv_cult$m3_beats_m2),2)*100`% and the `null` model in `r round(mean(cv_cult$m3_beats_m1),2)*100`% of cases. Furthermore, the `culture` model outperformed the `null` model in `r round(mean(cv_cult$m2_beats_m1),2)*100` % of cases. This suggests that there were substantial absolute and developmental differences between cultural settings (see Figure \@ref(fig:figdev)A).

## Processing signatures

In the second analysis focused on signatures in the data that suggest that children in all cultural settings use similar cognitive processes to infer the location of the balloon. The first -- somewhat weaker -- signature was simply an increase in performance with age. The second -- stronger -- signature was a strong effect of target centrality. That is, the further away from the center the balloon lands, the more imprecise should participants be. This second signature is derived from the model by the cognitive model introduced in @prein2023variation. It predicts that trials in which the agent looks further away from the center (i.e. to the left or right side of the screen) result in higher levels of imprecision compared to trials in which the agent looks closer to the middle.

These prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual's level of uncertainty in vector estimation and is represented in the model as a normal distribution around the true gaze vector with a participant specific standard deviation. On each trial, the model assumes that the participant samples a vector from this distribution to infer the location of the balloon. The smaller the standard deviation -- -- the narrower the light beam --, the more precise on average the participant is. Now, when the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area. This is true no matter how wide the cone (how large the standard deviation) is. A formal definition of the model can be found below.

```{r}
# draws <- bind_rows(
# as_draws_df(bm_cult_age)%>%
#   select(contains("ageinyears"))%>%
#   select(contains("ageinyears"))%>%
#   select(-contains("sd"))%>%
#   select(-contains("cor"))%>%
#   pivot_longer(cols = contains("ageinyears]"), names_to = "community", values_to = "value")%>%
#   mutate(post = value + b_ageinyears,
#          community = str_remove(community, "r_community"),
#          community = str_remove(community, ",ageinyears]"),
#          community = str_remove(community, "\\["))%>%
#   select(community, post)%>%
#   mutate(type = "ageinyears"),
# 
# as_draws_df(bm_cult_age)%>%
#   select(contains("targetcentralityx"))%>%
#   select(contains("targetcentralityx"))%>%
#   select(-contains("sd"))%>%
#   select(-contains("cor"))%>%
#   select(-contains("subjid"))%>%
#   pivot_longer(cols = contains("targetcentralityx]"), names_to = "community", values_to = "value")%>%
#   mutate(post = value + b_targetcentralityx,
#          community = str_remove(community, "r_community"),
#          community = str_remove(community, ",targetcentralityx]"),
#          community = str_remove(community, "\\["))%>%
#   select(community, post)%>%
#   mutate(type = "targetcentrality")
# )%>%
#   mutate(community = str_replace(community, "\\.", " "),
#          community= str_remove(community,"\\."))
# 
# pdraws <- draws%>%
#   mutate(community = recode(community,
#                             "Auckland (NewZealand)" = "Auckland (New Zealand)",
#                             "BaYaka (Rep.Congo)" = "BaYaka (Rep. Congo)",
#                             "Bandongo (Rep.Congo)" = "Bandongo (Rep. Congo)",
#                             "Buenos Aires(Argentina)" = "Buenos Aires (Argentina)"))%>%
#   left_join(data%>%distinct(community, continent))%>%
#   mutate(continent = factor(continent, levels = c("Americas", "Africa", "Europe", "Asia", "Oceania"), ordered = T))
# 
# saveRDS(pdraws, "../saves/pdraws.rds")
# 
# post_agg <- pdraws%>%
#   group_by(community, type)%>%
#   summarise(mean = mean(post),
#             lci = hdi_lower(post),
#             uci = hdi_upper(post))
# 
# saveRDS(post_agg, "../saves/post_agg.rds")
```

```{r figpreds, include = T, fig.align = "center", fig.cap = "Posterior distributions of model estimates for the poredictor variables by cultural setting.", out.width="100%"}

# pdraws <- readRDS("../saves/pdraws.rds")%>%
#   mutate(type = recode(type,
#                        ageinyears = "Age",
#                        targetcentrality = "Target centrality"),
#          type = factor(type, levels = c("Age","Target centrality")))
# 
# 
# 
# ppred <- ggplot(pdraws, aes(x= post, y = community))+
#   geom_density_ridges(col = "white", alpha = .75, fill = "grey")+
#   geom_vline(xintercept = 0, lty = 3)+
#   facet_grid(~type, scales = "free", space = "free_y")+
#   #stat_pointinterval(size = 0.1)+
#   #xlim(-0.15, 0.1)+
#   labs(x = "Posterior model estimate", y= "")+
#   theme_few()+
#   theme( strip.text.y.right = element_text(angle = 0))
# 
# 
# ggsave(plot = ppred, "../visuals/pred.png", height = 6, width = 12, scale = 1)

knitr::include_graphics("../visuals/pred.png")

```

```{r}
fixef_cult_age <- readRDS("../saves/cult_age_fix.rds")
```

To test these predictions, we inspected the parameter estimates for age and target centrality in the "winning" `culture_age` model described in the previous sections. We found strong main effects of age ($\beta$ = `r fixef_cult_age%>%filter(term == "ageinyears")%>%pull(Estimate)`, SE = `r fixef_cult_age%>%filter(term == "ageinyears")%>%pull(Est.Error)`, 95% CrI = `r fixef_cult_age%>%filter(term == "ageinyears")%>%pull(Q2.5)` - `r fixef_cult_age%>%filter(term == "ageinyears")%>%pull(Q97.5)`) and target centrality ($\beta$ =`r fixef_cult_age%>%filter(term == "targetcentralityx")%>%pull(Estimate)`, SE = `r fixef_cult_age%>%filter(term == "targetcentralityx")%>%pull(Est.Error)`, 95% CrI = `r fixef_cult_age%>%filter(term == "targetcentralityx")%>%pull(Q2.5)` - `r fixef_cult_age%>%filter(term == "targetcentralityx")%>%pull(Q97.5)`). Figure \@ref(fig:figpreds) visualizes the posterior distributions for each cultural setting and shows that -- despite variation in strength -- there was a substantial effect of age and target centrality in all cultural settings.

Another way to look at these signatures is to visualize the group-level developmental trajectories and the aggregated performance at different areas of the screen. Figure \@ref(fig:figdev)A shows that children got more precise with age in all cultural settings. Figure \@ref(fig:figdev)B shows the u-shaped pattern predicted by the cognitive model: the closer the balloon lands to the center of the screen (bins 5 and 6) the more precise children are in all cultural settings.

```{r}
# devdata <- data%>%
#   mutate(click = abs(clickdistfromtargetcenterx))%>%
#   group_by(subjid, continent,community, ageinyears)%>%
#   summarise(mean_click = mean(click))%>%
#   ungroup()%>%
#   mutate(ageinyears = ageinyears - min(ageinyears))
# 
# bm_dev <- brm(mean_click ~ ageinyears + (ageinyears |community),
#                               data   = devdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95))
# 
# ndata_all <- devdata%>%
#   distinct(ageinyears)
# 
# post_all <- fitted(bm_dev, newdata = ndata_all, re_formula = NA)%>%
#   as_tibble() %>%
#   bind_cols(ndata_all)%>%
#   mutate(ageinyears = ageinyears + min(data$ageinyears))
# 
# ndata <- devdata%>%
#   distinct(continent, community, ageinyears)
# 
# 
# post <- fitted(bm_dev, newdata = ndata, re_formula =~(ageinyears|community))%>%
#   as_tibble() %>%
#   bind_cols(ndata)%>%
#   mutate(ageinyears = ageinyears + min(data$ageinyears))
```

```{r}
# pagg_dev <- ggplot()+
#   geom_point(data = devdata, aes(x = ageinyears+ min(data$ageinyears), y = mean_click, col = community), alpha = 0.25)+
#   geom_smooth(data = post_all, aes(x = ageinyears, y = Estimate, ymin = Q2.5, ymax = Q97.5), stat = "identity", alpha = 0.5, col = "#555555", size = 1.5)+
#   theme_few()+
#     scale_x_continuous(limits = c(2,10))+
#   scale_y_continuous(limits = c(0,1000))+
#   scale_fill_manual(name = "Community", values = colors, guide = "none")+
#   scale_color_manual(name = "Community", values = colors, guide = "none")+
#   labs(x = "Age", y="Imprecision")+
#   theme(legend.position = "right")
# 
# pind <- ggplot()+ 
#  geom_point(data = devdata, aes(x = ageinyears+ min(data$ageinyears), y = mean_click, col = community), alpha = 0.25)+
#   geom_smooth(data = post, aes(x = ageinyears, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =community, col = community), stat = "identity", alpha = .25)+
#   theme_few()+
#   scale_fill_manual(name = "Community", values = colors, guide = "none")+
#   scale_color_manual(name = "color", values = colors, guide = "none")+
#   labs(x = "Age", y="Imprecision")+
#   scale_x_continuous(limits = c(2,10))+
#   scale_y_continuous(limits = c(0,1000))+
#   theme(legend.position = c(0.7, 0.9), legend.background = element_blank(),axis.title.x = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())+
#   facet_wrap(~community, nrow = 3)

```

```{r}
# tc_ind<- data%>%
#   mutate(targetposition = factor(targetposition, levels = c(1:10)))%>%
#   group_by(community, subjid, targetposition)%>%
#   summarise(mean = mean(abs(clickdistfromtargetcenterx)))
# 
# tc_agg<- data%>%
#   mutate(targetposition = factor(targetposition, levels = c(1:10)))%>%
#   group_by(targetposition)%>%
#   tidyboot_mean(col= abs(clickdistfromtargetcenterx))
# 
# tc_com<- data%>%
#   mutate(targetposition = factor(targetposition, levels = c(1:10)))%>%
#   group_by(community, targetposition)%>%
#   tidyboot_mean(col= abs(clickdistfromtargetcenterx))
# 
# 
# 
# ptagg <- ggplot()+
#   geom_jitter(data = tc_ind, aes(x = targetposition, y = mean, col = community), alpha = .05, height = 0, width = 0.2)+
#   geom_pointrange(data =tc_agg, aes(x = targetposition, y = mean, ymin = ci_lower, ymax = ci_upper),pch = "-", col = "#555555", size = 3, stroke = 1.5, linewidth = 2)+
#   labs(x = "Target position (binned)", y = "Imprecision")+
#   scale_color_manual(name = "", guide = "none", values = colors)+
#   ylim(0,1000)+
#   theme_few()
# 
# 
# ptind <- ggplot()+
#   geom_pointrange(data =tc_com, aes(x = targetposition, y = mean, ymin = ci_lower, ymax = ci_upper, col = community),pch = "-", size = 1)+
#   labs(x = "", y = "")+
#   scale_color_manual(name = "", guide = "none", values = colors)+
#   ylim(0,1000)+
#   theme_few()+
#   theme(legend.position = c(0.7, 0.9), legend.background = element_blank(),axis.title.x = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())+
#   facet_wrap(~community, nrow = 3)
```

```{r figdev, include = T, fig.align = "center", fig.cap = "A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by screen section across and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.", out.width="100%"}

# pvis_pred <- plot_grid(
# plot_grid(
#   pagg_dev, pind, rel_widths = c(1,1.5), align = "h", axis = "bt",labels = c("A", "B")
# ),
# plot_grid(
#   ptagg, ptind, rel_widths = c(1,1.5), align = "h", axis = "bt", labels = c("C", "D")
# ), nrow = 2, align = "v"
# )
# 
# ggsave(plot = pvis_pred, "../visuals/pvis_pred.png", width = 12, height = 8, bg = "white", scale = 1.25)

knitr::include_graphics("../visuals/pvis_pred.png")

```

### Cognitive model

The u-shaped pattern reported above may also arise if participants did not process the agent's gaze at all but only selected the middle of the screen or selected random locations. When only selecting the middle, the further away the balloon lands from the center, the further away it is from the center and the participant's click. When randomly selecting locations, the maximal imprecision is smaller when the target lands in the center (max: half screen width) compared to when it lands on the edge (max: full screen width). As a consequence, both alternative accounts would  result in a u-shaped pattern and a strong effect of target centrality. They would not, however, predict a developmental improvement in performance. Nevertheless, we wanted to quantify the evidence in favor of the vector-based gaze estimation model compared to these alternative center bias and random guessing models. We also added a random guessing model to the comparison according to which participants also ignored the agents gaze but simply touched random locations on the screen.

The vector-based gaze estimation model has been described in detail in @prein2023variation. In brief, the it inversely models the process generating clicks based on observed eye movements. Formally, the model is defined as:

\begin{equation}
    P(\theta | x_c, \alpha_l, \alpha_r) \propto P(x_c | \alpha_l, \alpha_r, \theta)P(\theta)
\end{equation}

Here, $\theta$ represents an individual's cognitive ability to locate the focus of the agent's attention, $x_c$ represents the clicked coordinate, and $\alpha_l$ and $\alpha_r$ correspond to the left and right pupil angles (each defined as the angle between a line connecting the center of the eye to the pupil and a line extended vertically downward from the center of the eye).

The basic assumption in this model is that participants click on the screen location where they think the agent is looking. The true eye angles ($\alpha_l$ and $\alpha_r$) are not directly observable and are estimated with noise, yielding $\hat{\alpha_l}$ and $\hat{\alpha_r}$.

Each click $x_c$ implies a "matched pair" of estimated pupil angles $\hat{\alpha_l}$ and $\hat{\alpha_r}$, with the constraint that the lines extended along those two angles meet at the precise location of where the target is believed to be. As a consequence, we can rewrite the likelihood function of the model as:

\begin{equation}
P(x_c | \alpha_l, \alpha_r, \theta) \propto P(\hat{\alpha_l}, \hat{\alpha_r} | \alpha_l, \alpha_r, \theta) P(x_c)
\end{equation}

$P(x_c)$ is a prior over potential target locations. Because the target was last visible in the screen and because the agent was located in the center, we assumed that participants have an a priori expectation that the target will land close to the middle. We estimated the strength of this center bias (i.e., the standard deviation of a Normal distribution around the screen center) based on the data: $P(x_c) \sim \mathcal{N}(960, \sigma^p)$.

The primary inferential task for participants is therefore to estimate the pupil angles ($\hat{\alpha_l}$ and $\hat{\alpha_r}$), i.e., to sample from the term $P(\hat{\alpha_l}, \hat{\alpha_r} | \alpha_l, \alpha_r, \theta)$. Here, we assumed that the pair of estimated pupil angles were sampled from a probability distribution which is the product of two Normal distributions of equal variance, $\sigma_v$, centered on the true pupil angles:

\begin{equation}
P(\hat{\alpha_l}, \hat{\alpha_r} | \alpha_l, \alpha_r, \theta) \propto \phi(\hat{\alpha}_l ; \alpha_l, \sigma_v)\phi(\hat{\alpha}_r ; \alpha_r, \sigma_v),
\end{equation}

Here, $\sigma_v$ plays a crucial role in determining the accuracy with which participants estimate pupil angles. Smaller values of $\sigma_v$ result in a narrow distribution around the pupil angle, making clicks far away from the target less likely. Conversely, larger values for $\sigma_v$ lead to a wider distribution, making clicks far away from the target more likely. To circle back to the analogy introduced above, $\sigma_v$ corresponds to the width of the light beam. Thus, the goal of the model was to estimate participant-specific values for $\sigma_v$: $\sigma_{v_i}$. These participant specific values were sampled from a normal distribution centered around the value we expected for a child given their age ($age_i$): $\sigma_{v_i} \sim e^{\mathcal{N}(\sigma_{v_{age}}, \sigma^{\sigma_{v_{age}}}})$. (Values were exponentiated to always be positive). $\sigma_{v_{age}}$ in turn was defined via a simple regression model: $\sigma_{v_{age}} = \beta_0^{\sigma_{v_{age}}} + age_i \cdot \beta_1^{\sigma_{v_{age}}}$. Where $\beta_0^{\sigma_{v_{age}}}$ and $\beta_1^{\sigma_{v_{age}}}$ represent the population-level intercept and slope that together describe a curve that represents the average developmental pattern.

To summarize, the model assumes that participant's clicks are generated by a process that relies on noisy estimates of the agent's gaze direction. The precision, with which the gaze direction is estimated, varies between participants and increases with development.

The two alternative models assume that participants ignore the agent's gaze completely, instead they are assumed to follow simple heuristics. According to the center bias model they always try to click in the center of the screen: $P(x_c) \sim \mathcal{N}(960, 160)$. (960 is the x-coordinate of the center and 160 is the width of the balloon). According to the random guessing model, they randomly click coordinates on the screen: $P(x_c) \sim \mathcal{U}(0, 1920)$.

All models were run separately for each cultural setting. The code to run the models can be found in the associated online repository. We also refer to this source for information on the prior distributions for all model parameters.

We compared models based on the marginal likelihood of the data for each model, which represents the likelihood of the data while averaging over the prior distribution on parameters. The pair-wise ratio of marginal likelihoods for two models is known as the Bayes Factor. Bayes Factors are a quantitative measure of the predictive quality of a model, taking into account the possible values of the model parameters weighted by their prior probabilities. The incorporation of the prior distribution over parameters in the averaging process implicitly considers model complexity: models with more parameters typically exhibit broader prior distributions over parameter values and broader prior distribution can attenuate the potential gains in predictive accuracy that a model with more parameters might otherwise achieve [@lee2014bayesian].

Table \@ref(tab:bftable) shows the Bayes Factors in favor of the vector-based gaze estimation model for each cultural setting. The evidence in favor of the gaze model compared to the two alternative models was clear (all BF > 1 000 000 000 000 000). This result rules out that the u-shaped pattern in the data was the result of participants always clicking the center instead of following gaze.

```{r bftable, echo=F}
bf <- readRDS("../model/output/bf_comparison.rds")%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(community = recode(community,
                            nigeria = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            china = "Beijing (China)", 
                            india = "Pune (India)",
                            argentina = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)"))

kableExtra::kable(bf,
             col.names = c("Community", "log(BF\\textsubscript{gaze,center})", "log(BF\\textsubscript{gaze,guess})"),
             caption = "Log Bayes Factors in favor of the gaze model compared to two alternative models.",
             format="latex", booktabs=TRUE, escape = F
             )
```

## Experiential predictors

```{r}
# scdata <- mdata%>%
#   group_by(subjid, community, ageinyears, touchscreen, household, children, younger_children)%>%
#   summarise(mean_click = mean(click))%>%
#   ungroup()%>%
#   group_by(community)%>%
#   mutate(
#          household = scale(household),
#          children = scale(children),
#          younger_children = scale(younger_children)
#          )%>%
#   na.omit()
# 
# bm_pred_null <- brm(mean_click ~ ageinyears + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc <- brm(mean_click ~ ageinyears + touchscreen + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8))%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_hh <- brm(mean_click ~ ageinyears + touchscreen + household + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# 
# bm_pred_tc_c <- brm(mean_click ~ ageinyears + touchscreen + children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_yc <- brm(mean_click ~ ageinyears + touchscreen + younger_children + (ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
# 
# bm_pred_tc_rand <- brm(mean_click ~ ageinyears + touchscreen + (touchscreen + ageinyears |community),
#                               data   = scdata,
#                               family = lognormal(),
#                               warmup = 2000,
#                               iter   = 4000,
#                               chains = 4,
#                               cores  = 4,
#              control = list(adapt_delta = 0.95),
#              backend = "cmdstanr",
#              threads = threading(8)
#              )%>%
#   add_criterion(., c("waic","loo"))
```

```{r}
# model_comp <- loo_compare(bm_pred_null, bm_pred_tc, bm_pred_tc_hh, bm_pred_tc_yc, bm_pred_tc_c, bm_pred_tc_rand,
#             criterion = "waic"
#             )%>%
#   as_tibble(rownames = "model")%>%
#   left_join(model_weights(bm_pred_null, bm_pred_tc, bm_pred_tc_hh, bm_pred_tc_yc, bm_pred_tc_c, bm_pred_tc_rand, criterion = "waic")%>%
#               as_tibble(rownames = "model")
#   )%>%
#   mutate_if(is.numeric, round, 2)
# 
# saveRDS(model_comp, "../saves/model_comp.rds")

model_comp <- readRDS("../saves/model_comp.rds")
```

The final analysis focused on whether we could predict performance in the task by aspects of everyday experience. For the ease of model fitting, we aggregated the data for each participant so that models predicted the average imprecision across trials^[@prein2023variation showed that the performance mean is nearly perfectly correlated with $\sigma_v$. When re-running the winning model on the trial-by-trial data, the results were qualitatively and quantitatively identical.]. The aspects we could focus on were limited to the questions we asked in the questionnaire. That is, we analysed the effects of exposure to technology and household composition.

We asked about children's exposure to screens as well as touchscreens. These two variables were largely redundant (see Figure \@ref(fig:figpred1)B) and so we included only one of them in the model. We chose the availability of a touchscreen as a predictor because the task itself was presented on a touchscreen. We expected (and pre-registered) that children with exposure to touchscreens would perform better.

For household composition, we asked for the total number of people in the household, the number of children and the number of younger children. We standardized each predictor *within* each culture before fitting the models. Thus, the interpretation of the coefficient is the gain in precision for living e.g., in a larger household relative to other children from the same cultural settings.

We compared models in a Bayesian model comparison. Models were compared based on the difference in expected log pointwise predictive density (`ELPD`) computed via the widely applicable information criterion (`WAIC`) and the standard error of that difference (SE). As pre-registered, we fit a separate model for each aspect of household composition. We compared the following models.

`null`:

* `mean_imprecision ~ age + (age | culture)`

`touchscreen`:

* `mean_imprecision ~ touchscreen + age + (age | culture)`

`household`:

* `mean_imprecision ~ household + touchscreen + age + (age | culture)`

`children`:

* `mean_imprecision ~ children + touchscreen + age + (age | culture)`

`younger_children`:

* `mean_imprecision ~ younger_children + touchscreen + age + (age | culture)`

```{r predcomptable, echo=F}
pred_mocel_comp <- readRDS("../saves/model_comp.rds")%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  select(model, elpd_diff, se_diff, waic, se_waic, value)%>%
  mutate(model = recode(model, 
                        bm_pred_null = "null",
                        bm_pred_tc = "touchscreen",
                        bm_pred_tc_hh = "household",
                        bm_pred_tc_c = "children",
                        bm_pred_tc_yc = "younger children",
                        bm_pred_tc_rand = "touchscreen (by culture)"))

kableExtra::kable(pred_mocel_comp,
             col.names = c("Model","diff\\textsubscript{WAIC}", "diff\\textsubscript{SE}", "WAIC", "SE", "Weight"),
             caption = "Log Bayes Factors in favor of the gaze model compared to two alternative models.",
             format="latex", booktabs=TRUE, escape = F
             )

```

Table \@ref(tab:predcomptable) shows that the model predicting average imprecision by the availability of touchscreens performed best and better than any of the models including a predictor referring to household composition. However, the difference between the models was very small (`ELPD` < 1, see Table \@ref(tab:predcomptable)), suggesting that they are largely equivalent, and, for the sake of consistency, we therefore inspected the posterior estimates for household composition predictors nevertheless.

To test whether there was variation across cultural settings in the importance of the availability of touchscreens, we added a model to the comparison that allowed this predictors to vary by cultural setting:

`touchscreen (by culture)`:

* `mean_imprecision ~ touchscreen + age + (touchscreen + age | culture)`

```{r}
# pred_est <- fixef(bm_pred_tc)%>%
#   as_tibble(rownames = "term")%>%
#   mutate_if(is.numeric,round, 2)
# 
# saveRDS(pred_est, "../saves/pred_est.rds")

pred_est <- readRDS("../saves/pred_est.rds")
```

Table \@ref(tab:predcomptable) shows that this model preformed relatively poor, suggesting that there was little variation across cultural settings in how the availability of touchscreens influenced performance. 
When inspecting the estimates within the winning model, we saw that children with exposure to touchscreens performed better compared to children without exposure ($\beta$ = `r pred_est%>%filter(term == "touchscreen")%>%pull(Estimate)`, SE = `r pred_est%>%filter(term == "touchscreen")%>%pull(Est.Error)`, 95% CrI = `r pred_est%>%filter(term == "touchscreen")%>%pull(Q2.5)` - `r pred_est%>%filter(term == "touchscreen")%>%pull(Q97.5)`). Figure \@ref(fig:figtcagg) shows that there was a very strong correlation between touchscreen availability and performance on a cultural setting-level as well. To account for age differences between settings in this analysis, we used random intercept estimates for each setting from a model predicting performance by age (the `null` model mentioned above). Thus, the values on the y-axis in Figure \@ref(fig:figtcagg) represent the deviation from the grand intercept for each setting for age = 0 (i.e., the mean given the centering procedure we applied before fitting the model).

When inspecting the posterior estimates for the different predictors referring to household composition, the overall pattern was consistent with the interpretation that social interactions, in particular with older children and adults, support the development of gaze following. However, all estimates were very small and not reliably different from zero. We think the reason for this lack of a strong association is a lack of resolution: household composition is very far removed from the factors that previous work has suggested to be related to the development of gaze following in younger children, such as attachment quality or the importance of gaze in early communicative interactions [@astor2020social; @movellan2002development; @senju2015early]. We hope that future work can increase the resolution with which everyday experiences in children from diverse cultural settings are recorded to compare the drivers behind development as we observe it.

```{r figtcagg, include = T, fig.align = "center", fig.cap = "Correlation between performance in the gaze-following task and the exposure to touchscreens on a cultural setting-level. Realtive imprecision accounts for differences in age between settings in that it corresponds to the random intercept estimate for each setting in a model predicting performance by age.", out.width="50%"}
# ptc_com <- ranef(bm_pred_null)$community[, ,1]%>%as_tibble(rownames = "community")%>%
#   left_join(data%>%group_by(community)%>%summarise(touchscreen = mean(touchscreen, na.rm = T)))
# 
# saveRDS(ptc_com, "../saves/ptc_com.rds")

# ptc_com <- readRDS("../saves/ptc_com.rds")
# 
# pcomtc <- ggplot(ptc_com,aes(x = touchscreen, y = Estimate))+
#   stat_cor(aes(x = touchscreen, y = Estimate,label = paste(..r.label..)),inherit.aes = F,r.accuracy = 0.01, cor.coef.name = "r",label.y = 0.85)+
#   geom_point(alpha = .75, pch = 1)+
#   geom_smooth(method = "lm", col = "firebrick", alpha = .25)+
#   labs(x = "Touchscreen exposure", y = "Relative imprecision (age corrected)")+
#   theme_few()
# 
# ggsave(plot = pcomtc, "../visuals/pcomtc.png", height = 4, width = 5, scale = 1.2)

knitr::include_graphics("../visuals/pcomtc.png")
```

# Appendix

## Voice-over script

### Welcome (before the start of the task)

Hello! Great that you’re here. We’ll now play a balloon game. Can you see the children in the picture over there? We want to play together with the children using the balloon. We’ll now talk you through exactly what will happen.

### Fam 1

Look, a girl/boy is standing in the window. And can you see the balloon over there? The balloon always falls down and lands on the ground. And you have to find it! The girl/boy helps you and always looks at the balloon.

[balloon falls down]

Where is the balloon? Touch the balloon!

### Fam 2

Perfect, that was great! Now, we’ll continue playing. Can you see the girl/boy and the balloon again? The balloon will fall down again. This time, it will fall behind a hedge. And you have to find it! The girl/boy helps you and looks at the balloon.

[balloon falls down]

Where is the balloon? On the hedge, touch where the balloon is.

### Test

Nice, good job! Now, we’ll continue playing. There is the balloon, the girl/boy and the hedge. The hedge is growing a bit now.

The balloon is behind the hedge now. You can’t see it - but the girl/boy can! The balloon falls to the ground and you have to find it. Remember - the girl/boy always looks at the balloon!

### Goodbye (after the last test trial)

The children are super happy after playing. Thanks a lot for your help! See you soon and goodbye!

\newpage

# References